---
title: "Analysis of soil at Casa de las Aguilas"
author: "Facundo Muñoz"
date: "\\today"
output:
  pdf_document:
    fig_caption: true
    pandoc_args: [
      "--output", "../reports/Casa_de_las_Aguilas.pdf"
    ]
---

```{r setup-knitr, echo = FALSE}
library(knitr)
opts_knit$set(root.dir = "..")
opts_chunk$set(echo       = FALSE,
               message    = FALSE,
               warning    = FALSE,
               comment    = NA,
               fig.width  = 4,
               fig.height = 4,
               fig.caption = TRUE,
               cache      = FALSE)
```

```{r setup-ProjectTemplate, include = FALSE}
## This automatically loads packages, data and helper functions
## and perform munging operations as defined for the project
library(ProjectTemplate)
load.project()
```


```{r setup-ggplot}
# theme_set(theme_bw())
theme_set(theme_tufte())

## To reuse in most plots:
## - structures + border + observations
## - equal scales + no labs
geom_aguilas <- function(obs)
  list(geom_polygon(aes(long, lat, group = group),
                    data = shp.construcciones),
  # geom_path(aes(group = group), data = shp.area.trabajo),
  geom_path(aes(x, y), data = border_coord),
  geom_point(aes(x, y), data = obs, alpha = 0.5, size = 1),
  coord_equal(),
  labs(x = NULL, y = NULL))

```


Here we compare the outcome of a classical kriging against a cost-based kriging which takes into account the presence of a semi-barrier.

## Data description

Figures 1 and 2 display the raw data, and an exploratory smoothed surface.


## Data description

```{r data-summary, fig.height=3, fig.cap='Empirical distribution of observations.'}

obs.wide <- spread(obs, var, value)
str(obs.wide)
summary(obs.wide)

ggplot(transform(obs, value = factor(value)), aes(value)) +
  geom_histogram() +
  facet_wrap(~ var) +
  theme_bw()
```



```{r raw-data, fig.height = 6, fig.cap = 'Measurement locations and observed values'}
# plot(shp.area.trabajo)
# plot(shp.construcciones)

# shp.construcciones@data$id <- rownames(shp.construcciones@data)
# structures <- fortify(shp.construcciones, region = 'id')
# structures <- join(structures, shp.construcciones@data, by = 'id')

  
p.pc <- ggplot(shp.construcciones, aes(long, lat)) +
  geom_aguilas(obs.pc) +
  geom_point(aes(x, y, color = pc), data = obs.pc, size = 5, shape = 15) +
  scale_color_viridis()

p.po4 <- ggplot(shp.construcciones, aes(long, lat)) +
  geom_aguilas(obs.pc) +
  geom_point(aes(x, y, color = po4), data = obs.po4, size = 5, shape = 15) +
  scale_color_viridis()

grid.arrange(p.pc, p.po4)
```



```{r smoothed-data, fig.cap = 'Exploratory kernel smoothing of the measurements'}

## Relative dimensions of the region
y2x_factor <- 1/do.call('/', as.list(diff(t(bbox(shp.contorno)))))
dimyx <- round(128 * c(y2x_factor, 1))
## use area.trabajo better than border!
## No: owin must be a (one) polygon.
## Otherwise, we could use a binary mask.
## But I think it is unimportant

## Create a `spatstat`'s Point Pattern object from the coordinates.
dat.pc.ppp <- as.ppp(obs.pc, owin(poly = rev(border_coord)))
# plot(dat.pc.ppp)

smooth.pc.im <- Smooth(dat.pc.ppp, dimyx = dimyx, sigma = 1)
smooth.pc.dat <- as.data.frame(smooth.pc.im)

dat.po4.ppp <- as.ppp(obs.po4, owin(poly = rev(border_coord)))
# plot(dat.po4.ppp)

smooth.po4.im <- Smooth(dat.po4.ppp, dimyx = dimyx, sigma = 1)
smooth.po4.dat <- as.data.frame(smooth.po4.im)

plotdat <- 
  rbind(data.frame(smooth.pc.dat,
                   var = 'pc'),
        data.frame(smooth.po4.dat,
                   var = 'po4'))

ggplot(plotdat, aes(x, y)) +
  geom_raster(aes(fill = value)) +
  geom_aguilas(data.frame(obs.pc, var = 'pc')) +
  geom_aguilas(data.frame(obs.po4, var = 'po4')) +
  scale_fill_viridis(limits = c(1, 6)) +
  facet_wrap(~var)


```



```{r cost-surface}

## Cost surface
res <- 0.2
prediction_grid <- raster(shp.area.trabajo, resolution = res)
cost_surf <- rasterize(shp.area.trabajo, prediction_grid)
# plot(cost_surf)


# ## Points exterior to all of the solid structures
# cost_loc <- coordinates(cost_surf)
# structure.coord.list <-
#   lapply(shp.construcciones@polygons,
#          function(x) coordinates(x@Polygons[[1]]))
# outer.idx <- 
#   apply(sapply(structure.coord.list,
#                function(x) point.in.polygon(cost_loc[, 'x'],
#                                             cost_loc[, 'y'],
#                                             x[, 1],
#                                             x[, 2]) == 0),
#         1, all)
# 
# 
# cost_surf[idx & !outer.idx] <- 0   # High friction == low "conductivity"
# plot(cost_surf)

```


```{r prediction-locations}
loc <- coordinates(cost_surf)[!is.na(values(cost_surf)),]
```

For the kriging predictions (both Euclidean and cost-based), we are using a
prediction grid with a resolution of `r res`, clipped by the structure of the
temple.


\pagebreak

## Percent of Carbonates

### Euclidean kriging

It may make sense to use the *room* as a covariate in this model (Universal kriging).
For the moment, we just perform an ordinary kriging.


The variogram model is Matérn.
We choose to estimate the nugget effect, which may account for measurement error, for example.


```{r pc-euclidean-kriging, include = FALSE}

## Observations as geodata
obs.pc.gd <- as.geodata(obs.pc, data.col="pc")

## compute euclidean (only) variogram
vg.std <- variog(obs.pc.gd)

## fitting variogram models
vgmdl.std <- likfit(geodata = obs.pc.gd,
                    fix.nugget = FALSE,
                    fix.kappa = FALSE,
                    kappa = 0.51,
                    ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                    cov.model = "exponential",
                    lik.method = "REML")

## Fitted parameters
par.tab <- with(vgmdl.std,
                data.frame(Euclidean = c(beta,
                                         nugget,
                                         sigmasq,
                                         kappa,
                                         phi,
                                         practicalRange),
                           row.names = c("Intercept",
                                         "Nugget",
                                         "Partial sill",
                                         "kappa",
                                         "phi",
                                         "Pract. range")))
  # , Cost_based=c(round(vgmdl.dmat$tausq,2),round(vgmdl.dmat$sigmasq,2),round(vgmdl.dmat$phi,0))


# Conventional Kriging, Euclidean distances
KC.std = krige.control(obj.model = vgmdl.std)
kriging.std <- krige.conv(obs.pc.gd, locations = loc, krige = KC.std)

```

```{r pc-euclidean-variogram-plot, fig.width = 6,fig.height = 3, fig.cap = 'Empirical Euclidean variogram and fitted model.'}

ggplot(vg.std) + geom_variogram(vgmdl.std) + xlim(c(0, 20))

```


```{r pc-euclidean-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Euclidean kriging prediction'}

ggplot(data.frame(loc, Prediction = kriging.std$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_aguilas(obs.pc) +
  scale_fill_viridis(limits = c(1, 6)) +
  scale_color_viridis(limits = c(1, 6))

```



### Cost-based kriging

Watch out! the cost surface can be derived either:
- from a SpatialPolygon of the working area
- from the SpatialPolygons of the border and of the inner structures

The results from both methods are not the same.
In the first case, the cost of non-conductive inner areas is NA, while in the second is 0.
This has an effect on one-pixel transitions (? this requires further inverstigation).


```{r pc-cost-based-distances}

ddm <- distmatGen(obs.pc[, c('x', 'y')], cost_surf)
ddm$loc <- ddm$loc[!is.na(values(cost_surf)),]
```

Some cost-based maps, for verification purposes.

```{r pc-some-cost-based-maps, fig.width = 10, fig.height = 7}
## Cost-based maps to each observation
cb.maps <-  cbind(loc,
                  ddm$loc) %>%
  as.data.frame() %>% 
  gather("Observation", "Distance", -(x:y))

idx <- c(10, 20, 90, 115, 280, 288)
obs.pc.idx <- cbind(obs.pc[idx, ],
                 Observation = factor(idx,
                                      levels = levels(cb.maps$Observation)))
ggplot(filter(cb.maps, Observation %in% idx), aes(x, y)) + 
  geom_tile(aes(fill = Distance, color = Distance)) +
  stat_contour(aes(z = Distance), color = 'lightgray') +
  geom_point(data = obs.pc.idx, col = 'red') +
  # geom_point(data = obs.pc[117, ], col = 'red') +
  # geom_point(data = filter(cb.maps, Observation %in% idx[1])[128, ], col = 'red') +
  # geom_polygon(aes(x, y), data = border_coord) +
  scale_fill_viridis() +
  scale_color_viridis() +
  geom_point(data = obs.pc, col = 'white', alpha = .5) +
  coord_fixed() +
  xlab('') + ylab('') +
  geom_aguilas(obs.pc) +
  facet_wrap(~Observation)

```


```{r pc-debug-locator, eval = FALSE}
# library(grid)
# tmp <- grid.locator('in')
# tmp.n <- as.numeric(tmp)
# tmp2.x <- as.numeric(convertX( unit(coordinates(cost_surf)[,'x'],'native'), 'in' ))
# tmp2.y <- as.numeric(convertY( unit(coordinates(cost_surf)[, 'y'],'native'), 'in' ))
# 
# w <- which.min( (tmp2.x-tmp.n[1])^2 + (tmp2.y-tmp.n[2])^2 )
# grid.text(w, tmp$x, tmp$y )
```



```{r pc-cost-based-kriging, include=FALSE}
## compute cost-based empirical variogram
vg.cst <- variog(obs.pc.gd, dists.mat = ddm$obs)

## fitting variogram models
vgmdl.cst <- likfit(geodata = obs.pc.gd,
                    fix.nugget = FALSE,
                    fix.kappa = FALSE,
                    kappa = 0.51,
                    ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                    cov.model = "exponential",
                    lik.method = "REML",
                    dists.mat = ddm$obs)

## Fitted parameters
par.tab$'Cost-based' <- with(vgmdl.cst,
                             c(beta,
                               nugget,
                               sigmasq,
                               kappa,
                               phi,
                               practicalRange))



## With trend, Cost-based distances
KC.cost = krige.control(obj.model = vgmdl.cst)

kriging.cst <- krige.conv(obs.pc.gd,
                          locations    = loc,
                          krige        = KC.cost,
                          dd.dists.mat = ddm$obs,
                          dl.dists.mat = ddm$loc)

```



```{r pc-cost-based-variogram-plot, fig.width = 6, fig.height = 3, fig.cap = 'Empirical Cost-based variogram and fitted model.'}

ggplot(vg.cst) + geom_variogram(vgmdl.cst) + xlim(c(0, 20))

```


```{r pc-cost-based-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Cost-based kriging prediction'}

ggplot(data.frame(loc, Prediction = kriging.cst$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_aguilas(obs.pc) +
  scale_fill_viridis(limits = c(0.99, 6.01)) +
  scale_color_viridis(limits = c(0.99, 6.01))

```



## Comparison of method outcomes

```{r pc-parameter-table}
kable(par.tab, digits = 2)
```


```{r pc-compare-variogram-plot, fig.width = 6, fig.cap = 'Empirical variograms and fitted models by method.'}

vg.both <- rbind(data.frame(vg.std[1:3],
                            method = 'Euclidean'),
                 data.frame(vg.cst[1:3],
                            method = 'Cost-based'))

names(vg.both) <- c('distance', 'semivariance', 'n', 'method')

ggplot(vg.both, aes(distance, semivariance)) +
  geom_point(aes(size = n)) + 
  expand_limits(y=0) +
  geom_variogram(vgmdl.std, method = 'Euclidean') +
  geom_variogram(vgmdl.cst, method = 'Cost-based') +
  xlim(c(0, 20)) +
  facet_grid(method ~.) +
  theme_bw()

```


The estimated variogram models are very similar in this case, with log-likelihoods of `r vgmdl.std$loglik` and `r vgmdl.cst$loglik` respectively.
This yields very similar kriging predictions as well.


```{r pc-cost-based-vs-euclidean-prediction}

data.frame(Euclidean = kriging.std$predict,
           Cost_based = kriging.cst$predict) %>% 
  ggplot(aes(Euclidean, Cost_based)) +
  geom_point() +
  geom_abline(int = 0, sl = 1, col = 'darkgray') +
  xlim(1, 6) + ylim(1, 6)
```



```{r pc-compare-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Comparison of Kriging estimates.'}

res.df <- data.frame(loc, 
                     method = c(rep('Classical', nrow(loc)),
                                rep('Cost-based', nrow(loc)),
                                rep('Difference', nrow(loc))),
                     Prediction = c(kriging.std$predict,
                                    kriging.cst$predict,
                                    kriging.cst$predict-kriging.std$predict))

pred.comp <- 
  ggplot(filter(res.df, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_aguilas(obs.pc) +
  scale_fill_viridis(limits = c(0.99, 6.01)) +
  scale_color_viridis(limits = c(0.99, 6.01)) +
  facet_wrap(~method) 

print(pred.comp)
```



```{r pc-compare-kriging-errors, fig.width=10, fig.height=7, fig.cap = 'Comparison of Kriging estimates.'}

err.df <- data.frame(
  loc, 
  method = c(rep('Classical', nrow(loc)),
             rep('Cost-based', nrow(loc)),
             rep('Difference', nrow(loc))),
  `Prediction error` = c(sqrt(kriging.std$krige.var),
                         sqrt(kriging.cst$krige.var),
                         sqrt(kriging.cst$krige.var)-sqrt(kriging.std$krige.var))
)

err.comp <- 
  ggplot(filter(err.df, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction.error, colour = Prediction.error)) +
  stat_contour(aes(z = Prediction.error), color = 'lightgray') +
  geom_aguilas(obs.pc) +
  scale_fill_viridis() +
  scale_color_viridis() +
  facet_wrap(~method) 

print(err.comp)
```


```{r pc-map-kriging-difference, fig.width=6, fig.height=7, fig.cap='Difference between the cost-based and the Euclidean predictions'}

pred.dif <- 
  ggplot(filter(res.df, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_aguilas(obs.pc) +
  scale_fill_viridis() +
  scale_color_viridis() +
  facet_wrap(~method)

err.dif <- 
  ggplot(filter(err.df, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction.error, colour = Prediction.error)) +
  stat_contour(aes(z = Prediction.error), color = 'lightgray') +
  geom_aguilas(obs.pc) +
  scale_fill_viridis() +
  scale_color_viridis() +
  facet_wrap(~method)

grid.arrange(pred.dif, err.dif)
```





\pagebreak

## Percent of Phosphates

### Euclidean kriging

It may make sense to use the *room* as a covariate in this model (Universal kriging).
For the moment, we just perform an ordinary kriging.


The variogram model is Matérn.
We choose to estimate the nugget effect, which may account for measurement error, for example.



```{r po4-euclidean-kriging, include = FALSE}
## Observations as geodata
obs.po4.gd <- as.geodata(obs.po4, data.col="po4")

## compute euclidean (only) variogram
vg.std <- variog(obs.po4.gd)

## fitting variogram models
vgmdl.std <- likfit(geodata = obs.po4.gd,
                    fix.nugget = FALSE,
                    fix.kappa = FALSE,
                    kappa = 0.51,
                    ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                    cov.model = "exponential",
                    lik.method = "REML")

## Fitted parameters
par.tab <- with(vgmdl.std,
                data.frame(Euclidean = c(beta,
                                         nugget,
                                         sigmasq,
                                         kappa,
                                         phi,
                                         practicalRange),
                           row.names = c("Intercept",
                                         "Nugget",
                                         "Partial sill",
                                         "kappa",
                                         "phi",
                                         "Pract. range")))
  # , Cost_based=c(round(vgmdl.dmat$tausq,2),round(vgmdl.dmat$sigmasq,2),round(vgmdl.dmat$phi,0))


# Conventional Kriging, Euclidean distances
KC.std = krige.control(obj.model = vgmdl.std)
kriging.std <- krige.conv(obs.po4.gd, locations = loc, krige = KC.std)

```

```{r po4-euclidean-variogram-plot, fig.width = 6,fig.height = 3, fig.cap = 'Empirical Euclidean variogram and fitted model.'}

ggplot(vg.std) + geom_variogram(vgmdl.std) + xlim(c(0, 20))

```


```{r po4-euclidean-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Euclidean kriging prediction'}

ggplot(data.frame(loc, Prediction = kriging.std$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_aguilas(obs.po4) +
  scale_fill_viridis(limits = c(1, 6)) +
  scale_color_viridis(limits = c(1, 6)) 
```



### Cost-based kriging

Watch out! the cost surface can be derived either:
- from a SpatialPolygon of the working area
- from the SpatialPolygons of the border and of the inner structures

The results from both methods are not the same.
In the first case, the cost of non-conductive inner areas is NA, while in the second is 0.
This has an effect on one-pixel transitions (? this requires further inverstigation).


```{r po4-cost-based-distances}

ddm <- distmatGen(obs.po4[, c('x', 'y')], cost_surf)
ddm$loc <- ddm$loc[!is.na(values(cost_surf)),]
```


```{r po4-cost-based-kriging, include=FALSE}

## compute cost-based empirical variogram
vg.cst <- variog(obs.po4.gd, dists.mat = ddm$obs)

## fitting variogram models
vgmdl.cst <- likfit(geodata = obs.po4.gd,
                    fix.nugget = FALSE,
                    fix.kappa = FALSE,
                    kappa = 0.51,
                    ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                    cov.model = "exponential",
                    lik.method = "REML",
                    dists.mat = ddm$obs)

## Fitted parameters
par.tab$'Cost-based' <- with(vgmdl.cst,
                             c(beta,
                               nugget,
                               sigmasq,
                               kappa,
                               phi,
                               practicalRange))



## With trend, Cost-based distances
KC.cost = krige.control(obj.model = vgmdl.cst)

kriging.cst <- krige.conv(obs.po4.gd,
                          locations    = loc,
                          krige        = KC.cost,
                          dd.dists.mat = ddm$obs,
                          dl.dists.mat = ddm$loc)

```



```{r po4-cost-based-variogram-plot, fig.width = 6, fig.height = 3, fig.cap = 'Empirical Cost-based variogram and fitted model.'}

ggplot(vg.cst) + geom_variogram(vgmdl.cst, max.dist = 20) + xlim(c(0, 20))

```


```{r po4-cost-based-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Cost-based kriging prediction'}

ggplot(data.frame(loc, Prediction = kriging.cst$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_aguilas(obs.po4) +
  scale_fill_viridis(limits = c(0.99, 6.01)) +
  scale_color_viridis(limits = c(0.99, 6.01))

```



## Comparison of method outcomes

```{r po4-parameter-table}
kable(par.tab, digits = 2)
```


```{r po4-compare-variogram-plot, fig.width = 6, fig.cap = 'Empirical variograms and fitted models by method.'}

vg.both <- rbind(data.frame(vg.std[1:3],
                            method = 'Euclidean'),
                 data.frame(vg.cst[1:3],
                            method = 'Cost-based'))

names(vg.both) <- c('distance', 'semivariance', 'n', 'method')

ggplot(vg.both, aes(distance, semivariance)) +
  geom_point(aes(size = n)) + 
  expand_limits(y=0) +
  geom_variogram(vgmdl.std, max.dist = 20, method = 'Euclidean') +
  geom_variogram(vgmdl.cst, max.dist = 20, method = 'Cost-based') +
  xlim(c(0, 20)) +
  facet_grid(method ~.) +
  theme_bw()

```


The estimated variogram models are very similar in this case, with log-likelihoods of `r vgmdl.std$loglik` and `r vgmdl.cst$loglik` respectively.
This yields very similar kriging predictions as well.

```{r po4-cost-based-vs-euclidean-prediction}

data.frame(Euclidean = kriging.std$predict,
           Cost_based = kriging.cst$predict) %>% 
  ggplot(aes(Euclidean, Cost_based)) +
  geom_point() +
  geom_abline(int = 0, sl = 1, col = 'darkgray') +
  xlim(1, 6) + ylim(1, 6)
```


```{r po4-compare-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Comparison of Kriging estimates.'}

res.df <- data.frame(loc, 
                     method = c(rep('Classical', nrow(loc)),
                                rep('Cost-based', nrow(loc)),
                                rep('Difference', nrow(loc))),
                     Prediction = c(kriging.std$predict,
                                    kriging.cst$predict,
                                    kriging.cst$predict-kriging.std$predict))

pred.comp <- 
  ggplot(filter(res.df, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_aguilas(obs.po4) +
  scale_fill_viridis(limits = c(0.99, 6.01)) +
  scale_color_viridis(limits = c(0.99, 6.01)) +
  facet_wrap(~method) 

print(pred.comp)
```



```{r po4-compare-kriging-errors, fig.width=10, fig.height=7, fig.cap = 'Comparison of Kriging estimates.'}

err.df <- data.frame(
  loc, 
  method = c(rep('Classical', nrow(loc)),
             rep('Cost-based', nrow(loc)),
             rep('Difference', nrow(loc))),
  `Prediction error` = c(sqrt(kriging.std$krige.var),
                         sqrt(kriging.cst$krige.var),
                         sqrt(kriging.cst$krige.var)-sqrt(kriging.std$krige.var))
)

err.comp <- 
  ggplot(filter(err.df, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction.error, colour = Prediction.error)) +
  stat_contour(aes(z = Prediction.error), color = 'lightgray') +
  geom_aguilas(obs.po4) +
  scale_fill_viridis() +
  scale_color_viridis() +
  facet_wrap(~method) 

print(err.comp)
```



```{r po4-map-kriging-difference, fig.width=6, fig.height=7, fig.cap='Difference between the cost-based and the Euclidean predictions'}

pred.dif <- 
  ggplot(filter(res.df, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_aguilas(obs.po4) +
  scale_fill_viridis() +
  scale_color_viridis() +
  facet_wrap(~method)

err.dif <- 
  ggplot(filter(err.df, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction.error, colour = Prediction.error)) +
  stat_contour(aes(z = Prediction.error), color = 'lightgray') +
  geom_aguilas(obs.po4) +
  scale_fill_viridis() +
  scale_color_viridis() +
  facet_wrap(~method)

grid.arrange(pred.dif, err.dif)
```



