---
title: "Analysis of soil at India"
author: "Facundo Muñoz"
date: "\\today"
output:
  pdf_document:
    toc: true
    toc_depth: 1
    number_sections: true
    fig_caption: true
    keep_tex: true
    pandoc_args: [
      "--output", "../reports/India.pdf"
    ]
---

\begin{flushright}
\texttt{geoRcb} package v.`r packageVersion('geoRcb')`
\end{flushright}

```{r setup-knitr, echo = FALSE}
library(knitr)
opts_knit$set(root.dir = "..")
opts_chunk$set(echo       = FALSE,
               message    = FALSE,
               warning    = FALSE,
               comment    = NA,
               fig.width  = 4,
               fig.height = 4,
               fig.caption = TRUE,
               cache      = FALSE)
```

```{r setup-ProjectTemplate, include = FALSE}
## This automatically loads packages, data and helper functions
## and perform munging operations as defined for the project
library(ProjectTemplate)
load.project()
```


```{r setup-ggplot}
# theme_set(theme_bw())
theme_set(theme_tufte())

## To reuse in most plots:
## - structures + border + observations
## - equal scales + no labs
geom_india <- function()
  list(geom_polygon(aes(long, lat, group = group),
                    data = shp.walls),
  geom_point(aes(x, y), data = dataset.Ca.Cu, alpha = 0.5, size = 1),
  coord_equal(),
  labs(x = NULL, y = NULL))

```


Here we compare the outcome of a classical kriging against a cost-based kriging which takes into account the presence of a barrier.

# Data description

```{r data-summary}
str(dataset.Ca.Cu)
summary(dataset.Ca.Cu)

ggplot(obs.india, aes(value)) +
  geom_histogram() +
  facet_wrap(~ variable, scales = 'free_x') 
```


Figures 1 and 2 display the raw data, and an exploratory smoothed surface.

```{r raw-data, fig.cap = 'Measurement locations and observed values'}
# plot(shp.area.trabajo)
# plot(shp.construcciones)

# shp.construcciones@data$id <- rownames(shp.construcciones@data)
# structures <- fortify(shp.construcciones, region = 'id')
# structures <- join(structures, shp.construcciones@data, by = 'id')

  
p.Ca <- ggplot(shp.walls, aes(long, lat)) +
  geom_india() +
  geom_point(aes(x, y, color = Ca), data = dataset.Ca.Cu, size = 5, shape = 15) +
  scale_color_viridis() 

p.Cu <- ggplot(shp.walls, aes(long, lat)) +
  geom_india() +
  geom_point(aes(x, y, color = Cu), data = dataset.Ca.Cu, size = 5, shape = 15) +
  scale_color_viridis()

grid.arrange(p.Ca, p.Cu)
```



```{r smoothed-data, fig.cap = 'Exploratory kernel smoothing of the measurements'}
## Create a `spatstat`'s Point Pattern object from the coordinates.
dat.ppp <- as.ppp(dataset.Ca.Cu[, -1],
                  owin(xrange = extent.india[1:2],
                       yrange = extent.india[3:4]))
# plot(dat.ppp)

## Relative dimensions of the region
y2x_factor <- 1/do.call('/', as.list(diff(t(bbox(extent.india)))))
dimyx <- round(128 * c(y2x_factor, 1))

smooth.im <- Smooth(dat.ppp, dimyx = dimyx, sigma = 1)
# plot(smooth.im)
smooth.dat <- as.data.frame(smooth.im)

p.Ca <- ggplot(smooth.dat, aes(Ca.x, Ca.y)) +
  geom_raster(aes(fill = Ca.value)) +
  geom_india() +
  scale_fill_viridis()

p.Cu <- ggplot(smooth.dat, aes(Cu.x, Cu.y)) +
  geom_raster(aes(fill = Cu.value)) +
  geom_india() +
  scale_fill_viridis()

grid.arrange(p.Ca, p.Cu)
```


\clearpage

# Cost-based distances

```{r setup-cost-surface}

## Cost surface
res <- 0.05
prediction_grid <- raster(extent.india, resolution = res)
cond_surf <- rasterize(shp.walls, prediction_grid, field = 0, background = 1)
# plot(cond_surf)


# ## Points exterior to all of the solid structures
# cost_loc <- coordinates(cond_surf)
# structure.coord.list <-
#   lapply(shp.construcciones@polygons,
#          function(x) coordinates(x@Polygons[[1]]))
# outer.idx <- 
#   apply(sapply(structure.coord.list,
#                function(x) point.in.polygon(cost_loc[, 'x'],
#                                             cost_loc[, 'y'],
#                                             x[, 1],
#                                             x[, 2]) == 0),
#         1, all)
# 
# 
# cond_surf[idx & !outer.idx] <- 0   # High friction == low "conductivity"
# plot(cond_surf)

```


```{r setup-obs-and-prediction-locations}
obs.gd <- as.geodata(dataset.Ca.Cu[, -1], data.col=c('Ca', 'Cu'))
loc <- coordinates(cond_surf)
```


```{r setup-cost-based-distances}

ddm <- distmatGen(dataset.Ca.Cu[, c('x', 'y')], cond_surf)
ddm$loc <- ddm$loc[!is.na(values(cond_surf)),]

```

Here we set up the cost-based surface, and compute some cost-based maps, for verifications purposes.

```{r setup-some-cost-based-maps, fig.width = 10, fig.height = 7, fig.cap = 'Some cost-based maps to selected observations.'}
## Cost-based maps to each observation
cb.maps <-  cbind(loc,
                  ddm$loc) %>%
  as.data.frame() %>% 
  gather("Observation", "Distance", -(x:y))

idx <- c(4, 30, 43, 69)
obs.idx <- cbind(dataset.Ca.Cu[idx, ],
                 Observation = factor(levels(cb.maps$Observation)[idx],
                                      levels = levels(cb.maps$Observation)))
ggplot(filter(cb.maps, Observation %in% levels(cb.maps$Observation)[idx]), aes(x, y)) + 
  geom_tile(aes(fill = Distance, color = Distance)) +
  stat_contour(aes(z = Distance), color = 'lightgray') +
  geom_point(data = obs.idx, col = 'red') +
  scale_fill_viridis() +
  scale_color_viridis() +
  geom_point(data = obs.india, col = 'white', alpha = .5) +
  coord_fixed() +
  xlab('') + ylab('') +
  geom_india() +
  facet_wrap(~Observation)

```






\clearpage

# Analysis of Calcium


## Euclidean kriging


The variogram model is Exponential.
We choose to estimate the nugget effect, which may account for measurement error, for example.



```{r Ca-euclidean-kriging, include = FALSE}

## compute euclidean (only) variogram
vg.std.Ca <- variog(obs.gd, data = obs.gd$data[, 'Ca'])

## fitting variogram models
vgmdl.std.Ca <- likfit(geodata = obs.gd,
                       data = obs.gd$data[, 'Ca'],
                       fix.nugget = FALSE,
                       fix.kappa = FALSE,
                       # kappa = 0.51,  # If 0.5 then cov.model changes to exponential
                       ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                       cov.model = "exponential",
                       lik.method = "REML")

## Fitted parameters
par.tab <- data.frame(
  Euclidean = with(vgmdl.std.Ca,
                   c(beta,
                     nugget,
                     sigmasq,
                     # kappa,
                     phi,
                     practicalRange,
                     loglik)),
  row.names = c("Intercept",
                "Nugget",
                "Partial sill",
                # "kappa",
                "phi",
                "Pract. range",
                "Log-likelihood"))


# Conventional Kriging, Euclidean distances
KC.std.Ca = krige.control(obj.model = vgmdl.std.Ca)
kriging.std.Ca <- krige.conv(obs.gd,
                             data = obs.gd$data[, 'Ca'],
                             locations = loc,
                             krige = KC.std.Ca)

```

```{r Ca-euclidean-variogram-plot, fig.width = 6,fig.height = 3, fig.cap = 'Empirical variogram and fitted model.'}

ggplot(vg.std.Ca) + 
  geom_variogram(vgmdl.std.Ca) + 
  ylab('semivariance Ca')

```


```{r Ca-euclidean-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Euclidean kriging prediction'}

p.Ca <- ggplot(data.frame(loc, Prediction = kriging.std.Ca$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis()

p.Ca

```



## Cost-based kriging



```{r Ca-cost-based-kriging, include=FALSE}


## compute cost-based empirical variogram
vg.cst.Ca <- variog(obs.gd, data = obs.gd$data[, 'Ca'], dists.mat = ddm$obs)

## fitting variogram models
vgmdl.cst.Ca <- likfit(geodata = obs.gd,
                       data = obs.gd$data[, 'Ca'],
                       fix.nugget = FALSE,
                       fix.kappa = FALSE,
                       kappa = 0.51,  # If 0.5 then cov.model changes to exponential
                       ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                       cov.model = "exponential",
                       lik.method = "REML",
                       dists.mat = ddm$obs)

## Fitted parameters
par.tab <- cbind(par.tab,
                 data.frame(
                   Cost_based = with(vgmdl.cst.Ca,
                                     c(beta,
                                       nugget,
                                       sigmasq,
                                       # kappa,
                                       phi,
                                       practicalRange,
                                       loglik))))


# Conventional Kriging, Euclidean distances
KC.cst.Ca = krige.control(obj.model = vgmdl.cst.Ca)
kriging.cst.Ca <- krige.conv(obs.gd,
                             data = obs.gd$data[, 'Ca'],
                             locations = loc,
                             krige = KC.cst.Ca,
                             dd.dists.mat = ddm$obs,
                             dl.dists.mat = ddm$loc)

```



```{r Ca-cost-based-variogram-plot, fig.width = 6, fig.height = 3, fig.cap = 'Empirical cost-based variogram and fitted model.'}

ggplot(vg.cst.Ca) +
  geom_variogram(vgmdl.cst.Ca) +
  ylab('semivariance Ca')

```


```{r Ca-cost-based-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Cost-based kriging prediction'}

p.Ca <- ggplot(data.frame(loc, Prediction = kriging.cst.Ca$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis()

p.Ca

```



## Comparison of method outcomes

```{r Ca-parameter-table}
kable(par.tab, digits = 2)
```


```{r Ca-compare-variogram-plot, fig.width = 6, fig.height = 4, fig.cap = 'Empirical variogram and fitted models by method for Calcium.'}

vg.both <- rbind(data.frame(vg.std.Ca[1:3],
                            variable = 'Ca',
                            method = 'Euclidean'),
                 data.frame(vg.cst.Ca[1:3],
                            variable = 'Ca',
                            method = 'Cost-based'))

names(vg.both) <- c('distance', 'semivariance', 'n', 'variable', 'method')

p.Ca <- ggplot(filter(vg.both, variable == 'Ca'),
               aes(distance, semivariance)) +
  geom_point(aes(size = n)) + 
  expand_limits(y=0) +
  geom_variogram(vgmdl.std.Ca, variable = 'Ca', method = 'Euclidean') +
  geom_variogram(vgmdl.cst.Ca, variable = 'Ca', method = 'Cost-based') +
  facet_grid(method ~ .) +
  theme_bw()

p.Ca

```




```{r Ca-compare-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Comparison of Kriging estimates.'}

res.df.Ca <- data.frame(loc, 
                        method = c(rep('Classical', nrow(loc)),
                                   rep('Cost-based', nrow(loc)),
                                   rep('Difference', nrow(loc))),
                        Prediction = c(kriging.std.Ca$predict,
                                          kriging.cst.Ca$predict,
                                          kriging.std.Ca$predict-kriging.cst.Ca$predict))

p.Ca <- 
  ggplot(filter(res.df.Ca, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .2, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis() +
  facet_wrap(~method, ncol = 1) 


p.Ca
```


```{r Ca-cost-based-vs-euclidean-prediction, fig.cap = 'Pointwise comparison of predictions by method.'}

p.Ca <- data.frame(Euclidean  = kriging.std.Ca$predict,
                   Cost_based = kriging.cst.Ca$predict,
                   variable   = 'Ca') %>% 
  ggplot(aes(Euclidean, Cost_based)) +
  geom_point() +
  geom_abline(int = 0, sl = 1, col = 'darkgray') +
  theme_bw()


p.Ca

```

In the scatter plot, the horizontal patterns correspond to predictions on observed values.
Otherwise, the differences are negligeable.

```{r Ca-map-kriging-difference, fig.width=10, fig.height=6, fig.cap='Difference between the Euclidean and the cost-based predictions.'}

pred.dif.Ca <- 
  ggplot(filter(res.df.Ca, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .2, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Difference in\npredictions') +
  scale_color_viridis(name = 'Difference in\npredictions')

pred.dif.Ca
# grid.arrange(pred.comp, pred.dif)
```





```{r Ca-compare-kriging-errors, fig.width=10, fig.height=7, fig.cap = 'Comparison of prediction error by method.'}

res.var.Ca <- data.frame(loc, 
                        method = c(rep('Classical', nrow(loc)),
                                   rep('Cost-based', nrow(loc)),
                                   rep('Difference', nrow(loc))),
                        Prediction = c(kriging.std.Ca$krige.var,
                                          kriging.cst.Ca$krige.var,
                                          kriging.std.Ca$krige.var-kriging.cst.Ca$krige.var))

ggplot(filter(res.var.Ca, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .2, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Kriging\nvariance') +
  scale_color_viridis(name = 'Kriging\nvariance') +
  facet_wrap(~method, ncol = 1) 

```


```{r Ca-map-kriging-difference-errors, fig.width=10, fig.height=6, fig.cap='Difference between the Euclidean and the cost-based prediction errors'}

ggplot(filter(res.var.Ca, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .2, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Difference in\nKriging variance') +
  scale_color_viridis(name = 'Difference in\nKriging variance')
```

Near the observations, the cost-based approach has a larger prediction error due to its increased estimation of the nugget (i.e. short-range variance).
In the main area, the prediction errors are practically the same with both approaches.
Behind the walls, the Euclidean prediction error is unrealistically low.


\clearpage

# Analysis of Copper


## Euclidean kriging


The variogram model is Exponential.
We choose to estimate the nugget effect, which may account for measurement error, for example.


```{r Cu-euclidean-kriging, include = FALSE}

## compute euclidean (only) variogram
vg.std.Cu <- variog(obs.gd, data = obs.gd$data[, 'Cu'])

## fitting variogram models
vgmdl.std.Cu <- likfit(geodata = obs.gd,
                       data = obs.gd$data[, 'Cu'],
                       fix.nugget = FALSE,
                       fix.kappa = FALSE,
                       kappa = 0.51,  # If 0.5 then cov.model changes to exponential
                       ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                       cov.model = "exponential",
                       lik.method = "REML")

## Fitted parameters
par.tab <- data.frame(Euclidean = with(vgmdl.std.Cu,
                                       c(beta,
                                         nugget,
                                         sigmasq,
                                         kappa,
                                         phi,
                                         practicalRange)),
                      row.names = c("Intercept",
                                    "Nugget",
                                    "Partial sill",
                                    "kappa",
                                    "phi",
                                    "Pract. range"))


# Conventional Kriging, Euclidean distances
KC.std.Cu = krige.control(obj.model = vgmdl.std.Cu)
kriging.std.Cu <- krige.conv(obs.gd,
                             data = obs.gd$data[, 'Cu'],
                             locations = loc,
                             krige = KC.std.Cu)

```

```{r Cu-euclidean-variogram-plot, fig.width = 6,fig.height = 3, fig.cap = 'Empirical variogram and fitted model for Copper.'}

ggplot(vg.std.Cu) +
  geom_variogram(vgmdl.std.Cu) + 
  ylab('semivariance Cu')

```


```{r Cu-euclidean-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Euclidean kriging prediction for Copper.'}

p.Cu <- ggplot(data.frame(loc, Prediction = kriging.std.Cu$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .5, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis()

p.Cu

```



## Cost-based kriging


```{r Cu-cost-based-kriging, include=FALSE}


## compute cost-based empirical variogram
vg.cst.Cu <- variog(obs.gd, data = obs.gd$data[, 'Cu'], dists.mat = ddm$obs)

## fitting variogram models
vgmdl.cst.Cu <- likfit(geodata = obs.gd,
                       data = obs.gd$data[, 'Cu'],
                       fix.nugget = FALSE,
                       fix.kappa = FALSE,
                       kappa = 0.51,  # If 0.5 then cov.model changes to exponential
                       ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                       cov.model = "exponential",
                       lik.method = "REML",
                       dists.mat = ddm$obs)

## Fitted parameters
par.tab <- cbind(par.tab,
                 data.frame(
                   Cost_based = with(vgmdl.cst.Cu,
                                     c(beta,
                                       nugget,
                                       sigmasq,
                                       kappa,
                                       phi,
                                       practicalRange))))


# Conventional Kriging, Euclidean distances
KC.cst.Cu = krige.control(obj.model = vgmdl.cst.Cu)
kriging.cst.Cu <- krige.conv(obs.gd,
                             data = obs.gd$data[, 'Cu'],
                             locations = loc,
                             krige = KC.cst.Cu,
                             dd.dists.mat = ddm$obs,
                             dl.dists.mat = ddm$loc)

```



```{r Cu-cost-based-variogram-plot, fig.width = 6, fig.height = 3, fig.cap = 'Empirical cost-based variogram and fitted model.'}

ggplot(vg.cst.Cu) +
  geom_variogram(vgmdl.cst.Cu) + 
  ylab('semivariance Cu')

```


```{r Cu-cost-based-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Cost-based kriging prediction'}

p.Cu <- ggplot(data.frame(loc, Prediction = kriging.cst.Cu$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .5, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis()

p.Cu

```



## Comparison of method outcomes

```{r Cu-parameter-table}
kable(par.tab, digits = 2)
```


```{r Cu-compare-variogram-plot, fig.width = 6, fig.height = 4, fig.cap = 'Empirical variogram and fitted models by method for Copper.'}

vg.both <- rbind(data.frame(vg.std.Cu[1:3],
                            variable = 'Cu',
                            method = 'Euclidean'),
                 data.frame(vg.cst.Cu[1:3],
                            variable = 'Cu',
                            method = 'Cost-based'))

names(vg.both) <- c('distance', 'semivariance', 'n', 'variable', 'method')

p.Cu <- ggplot(filter(vg.both, variable == 'Cu'),
               aes(distance, semivariance)) +
  geom_point(aes(size = n)) + 
  expand_limits(y=0) +
  geom_variogram(vgmdl.std.Cu, variable = 'Cu', method = 'Euclidean') +
  geom_variogram(vgmdl.cst.Cu, variable = 'Cu', method = 'Cost-based') +
  facet_grid(method ~ .) +
  theme_bw()

p.Cu

```




```{r Cu-compare-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Comparison of Kriging estimates.'}

res.df.Cu <- data.frame(loc, 
                        method = c(rep('Classical', nrow(loc)),
                                   rep('Cost-based', nrow(loc)),
                                   rep('Difference', nrow(loc))),
                        Prediction = c(kriging.std.Cu$predict,
                                       kriging.cst.Cu$predict,
                                       kriging.std.Cu$predict-kriging.cst.Cu$predict))

p.Cu <- 
  ggplot(filter(res.df.Cu, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .5, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis() +
  facet_wrap(~method, ncol = 1) 

p.Cu
```


```{r Cu-cost-based-vs-euclidean-prediction, fig.cap = 'Pointwise comparison of predictions by method.'}

p.Cu <- data.frame(Euclidean  = kriging.std.Cu$predict,
                   Cost_based = kriging.cst.Cu$predict,
                   variable   = 'Cu') %>% 
  ggplot(aes(Euclidean, Cost_based)) +
  geom_point() +
  geom_abline(int = 0, sl = 1, col = 'darkgray') +
  facet_grid(~variable) + 
  theme_bw()

p.Cu

```

In the scatter plot, the horizontal patterns correspond to predictions on observed values.
Otherwise, the differences are negligeable.

```{r Cu-map-kriging-difference, fig.width=10, fig.height=12, fig.cap='Difference between the Euclidean and the cost-based predictions.'}

pred.dif.Cu <- 
  ggplot(filter(res.df.Cu, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .5, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis('Difference in\npredictions') +
  scale_color_viridis('Difference in\npredictions') 

pred.dif.Cu
# grid.arrange(pred.comp, pred.dif)
```




```{r Cu-compare-kriging-errors, fig.width=10, fig.height=7, fig.cap = 'Comparison of prediction error by method.'}

res.var.Cu <- data.frame(loc, 
                        method = c(rep('Classical', nrow(loc)),
                                   rep('Cost-based', nrow(loc)),
                                   rep('Difference', nrow(loc))),
                        Prediction = c(kriging.std.Cu$krige.var,
                                          kriging.cst.Cu$krige.var,
                                          kriging.std.Cu$krige.var-kriging.cst.Cu$krige.var))

ggplot(filter(res.var.Cu, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Kriging\nvariance') +
  scale_color_viridis(name = 'Kriging\nvariance') +
  facet_wrap(~method, ncol = 1) 

```


```{r Cu-map-kriging-difference-errors, fig.width=10, fig.height=6, fig.cap='Difference between the Euclidean and the cost-based prediction errors'}

ggplot(filter(res.var.Cu, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Difference in\nKriging variance') +
  scale_color_viridis(name = 'Difference in\nKriging variance')
```

Near the observations, the cost-based approach has a larger prediction error due to its increased estimation of the nugget (i.e. short-range variance).
In the main area, the prediction errors are practically the same with both approaches.
Behind the walls, the Euclidean prediction error is unrealistically low.



\clearpage

# Conclusions

- Actually, the kriging model is not adjusting very well the tails of the data, which are heavier than expected.
  This happens both for the Euclidean and cost-based models. This means that none of both approaches will be really good predictors anyways.

```{r model-diagnostics}
## Not really a nice fit
hist(resid(vgmdl.std.Ca))
hist(resid(vgmdl.cst.Ca))

qqnorm(resid(vgmdl.std.Ca))
qqline(resid(vgmdl.std.Ca))

qqnorm(resid(vgmdl.cst.Ca))
qqline(resid(vgmdl.cst.Ca))

```

