---
title: "Analysis of soil at India"
author: "Facundo Muñoz"
date: "\\today"
output:
  pdf_document:
    toc: true
    toc_depth: 1
    number_sections: true
    fig_caption: true
    keep_tex: true
    pandoc_args: [
      "--output", "../reports/India.pdf"
    ]
---

\begin{flushright}
\texttt{geoRcb} package v.`r packageVersion('geoRcb')`
\end{flushright}

```{r setup-knitr, echo = FALSE}
library(knitr)
opts_knit$set(root.dir = "..")
opts_chunk$set(echo       = FALSE,
               message    = FALSE,
               warning    = FALSE,
               comment    = NA,
               fig.width  = 4,
               fig.height = 4,
               fig.caption = TRUE,
               cache      = TRUE)
```

```{r setup-ProjectTemplate, include = FALSE, cache = FALSE}
## This automatically loads packages, data and helper functions
## and perform munging operations as defined for the project
library(ProjectTemplate)
load.project()
```


```{r setup-ggplot, cache = FALSE}
# theme_set(theme_bw())
theme_set(theme_tufte())

## To reuse in most plots:
## - structures + border + observations
## - equal scales + no labs
geom_india <- function()
  list(geom_polygon(aes(long, lat, group = group),
                    data = shp.walls),
  geom_point(aes(x, y), data = dataset, alpha = 0.5, size = 1),
  coord_equal(),
  labs(x = NULL, y = NULL))

```


Here we compare the outcome of a classical kriging against a cost-based kriging which takes into account the presence of a barrier.

# Data description

```{r data-summary, fig.cap = 'Histograms of measured variables.', fig.width = 5, fig.height = 7}
str(dataset)
summary(dataset)

ggplot(obs.india, aes(value)) +
  geom_histogram() +
  facet_wrap(~ variable, scales = 'free_x', ncol = 2) 
```


Figures 1 and 2 display the raw data, and an exploratory smoothed surface.

```{r raw-data, fig.cap = 'Measurement locations and observed values', fig.width = 8, fig.height = 10}
# plot(shp.area.trabajo)
# plot(shp.construcciones)

# shp.construcciones@data$id <- rownames(shp.construcciones@data)
# structures <- fortify(shp.construcciones, region = 'id')
# structures <- join(structures, shp.construcciones@data, by = 'id')

plot_raw_variable <- function(var) {
  p <- ggplot(shp.walls, aes(long, lat)) +
    geom_india() +
    scale_color_viridis() 
  
  gp <- paste('geom_point(aes(x, y, color =', 
              parse(text= var), 
              '), data = dataset, size = 5, shape = 15)')

  p + eval(parse(text = gp))
}

p.list <- lapply(
  levels(obs.india$variable),
  plot_raw_variable
)

do.call('grid.arrange', c(p.list, list(ncol = 2)))
```



```{r smoothed-data, fig.cap = 'Exploratory kernel smoothing of the measurements', fig.width = 8, fig.height = 10}
## Create a `spatstat`'s Point Pattern object from the coordinates.
dat.ppp <- as.ppp(dataset[, -(1:3)],
                  owin(xrange = extent.india[1:2],
                       yrange = extent.india[3:4]))
# plot(dat.ppp)

## Relative dimensions of the region
y2x_factor <- 1/do.call('/', as.list(diff(t(bbox(extent.india)))))
dimyx <- round(128 * c(y2x_factor, 1))

smooth.im <- Smooth(dat.ppp, dimyx = dimyx, sigma = 1)
# plot(smooth.im)
smooth.dat <- as.data.frame(smooth.im)

plot_smoothed_variable <- function(var) {
  pt <- paste0('ggplot(smooth.dat, aes(', var, '.x, ', var, '.y))') 
  rt <- paste0('geom_raster(aes(fill = ', var, '.value))')
  
  eval(parse(text = pt)) + 
    eval(parse(text = rt)) +
    geom_india() +
    scale_fill_viridis() 
}

p.list <- lapply(
  levels(obs.india$variable),
  plot_smoothed_variable
)

do.call('grid.arrange', c(p.list, list(ncol = 2)))

```


\clearpage

# Cost-based distances

```{r setup-cost-surface}

## Cost surface
res <- 0.05
prediction_grid <- raster(extent.india, resolution = res)
cond_surf <- rasterize(shp.walls, prediction_grid, field = 0, background = 1)
# plot(cond_surf)


# ## Points exterior to all of the solid structures
# cost_loc <- coordinates(cond_surf)
# structure.coord.list <-
#   lapply(shp.construcciones@polygons,
#          function(x) coordinates(x@Polygons[[1]]))
# outer.idx <- 
#   apply(sapply(structure.coord.list,
#                function(x) point.in.polygon(cost_loc[, 'x'],
#                                             cost_loc[, 'y'],
#                                             x[, 1],
#                                             x[, 2]) == 0),
#         1, all)
# 
# 
# cond_surf[idx & !outer.idx] <- 0   # High friction == low "conductivity"
# plot(cond_surf)

```


```{r setup-obs-and-prediction-locations}
obs.gd <- as.geodata(dataset[, -(1:3)], 
                     data.col=levels(obs.india$variable))
loc <- coordinates(cond_surf)
```


```{r setup-cost-based-distances}

ddm <- distmatGen(dataset[, c('x', 'y')], cond_surf)
ddm$loc <- ddm$loc[!is.na(values(cond_surf)),]

## Remove prediction locations infinitely far from all points
## This is an artifact of the rasterization
idx <- which(apply(ddm$loc, 1,
             function(x) all(is.infinite(x))))

loc <- loc[-idx, ]
ddm$loc <- ddm$loc[-idx, ]

```

Here we set up the cost-based surface, and compute some cost-based maps, for verifications purposes.

```{r setup-some-cost-based-maps, fig.width = 10, fig.height = 7, fig.cap = 'Some cost-based maps to selected observations.'}
## Cost-based maps to each observation
cb.maps <-  cbind(loc,
                  ddm$loc) %>%
  as.data.frame() %>% 
  gather("Observation", "Distance", -(x:y))

idx <- c(4, 30, 43, 69)
obs.idx <- cbind(dataset[idx, ],
                 Observation = factor(levels(cb.maps$Observation)[idx],
                                      levels = levels(cb.maps$Observation)))
ggplot(filter(cb.maps, Observation %in% levels(cb.maps$Observation)[idx]), aes(x, y)) + 
  geom_tile(aes(fill = Distance, color = Distance)) +
  stat_contour(aes(z = Distance), binwidth = 1, color = 'lightgray') +
  geom_point(data = obs.idx, col = 'red') +
  scale_fill_viridis() +
  scale_color_viridis() +
  geom_point(data = obs.india, col = 'white', alpha = .5) +
  coord_fixed() +
  xlab('') + ylab('') +
  geom_india() +
  facet_wrap(~Observation)

```






\clearpage

# Analysis of Calcium

```{r Ca-fix-variable}
variable <- obs.gd$data[, 'Ca']
```


## Euclidean kriging


The variogram model is Exponential.
We choose to estimate the nugget effect, which may account for measurement error, for example.



```{r Ca-euclidean-kriging, include = FALSE}

## compute euclidean (only) variogram
vg.std.Ca <- variog(obs.gd, data = variable)

## fitting variogram models
vgmdl.std.Ca <- likfit(geodata = obs.gd,
                       data = variable,
                       fix.nugget = FALSE,
                       fix.kappa = FALSE,
                       # kappa = 0.51,  # If 0.5 then cov.model changes to exponential
                       ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                       cov.model = "exponential",
                       lik.method = "REML")

## Fitted parameters
par.tab <- data.frame(
  Euclidean = with(vgmdl.std.Ca,
                   c(beta,
                     nugget,
                     sigmasq,
                     # kappa,
                     phi,
                     practicalRange,
                     loglik)),
  row.names = c("Intercept",
                "Nugget",
                "Partial sill",
                # "kappa",
                "phi",
                "Pract. range",
                "Log-likelihood"))


# Conventional Kriging, Euclidean distances
KC.std.Ca = krige.control(obj.model = vgmdl.std.Ca)
kriging.std.Ca <- krige.conv(obs.gd,
                             data = variable,
                             locations = loc,
                             krige = KC.std.Ca)

```

```{r Ca-euclidean-variogram-plot, fig.width = 6,fig.height = 3, fig.cap = 'Empirical variogram and fitted model.'}

ggplot(vg.std.Ca) + 
  geom_variogram(vgmdl.std.Ca) + 
  ylab('semivariance Ca')

```


```{r Ca-euclidean-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Euclidean kriging prediction'}

p.Ca <- ggplot(data.frame(loc, Prediction = kriging.std.Ca$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis()

p.Ca

```



## Cost-based kriging



```{r Ca-cost-based-kriging, include=FALSE}


## compute cost-based empirical variogram
vg.cst.Ca <- variog(obs.gd, data = variable, dists.mat = ddm$obs)

## fitting variogram models
vgmdl.cst.Ca <- likfit(geodata = obs.gd,
                       data = variable,
                       fix.nugget = FALSE,
                       fix.kappa = FALSE,
                       kappa = 0.51,  # If 0.5 then cov.model changes to exponential
                       ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                       cov.model = "exponential",
                       lik.method = "REML",
                       dists.mat = ddm$obs)

## Fitted parameters
par.tab <- cbind(par.tab,
                 data.frame(
                   Cost_based = with(vgmdl.cst.Ca,
                                     c(beta,
                                       nugget,
                                       sigmasq,
                                       # kappa,
                                       phi,
                                       practicalRange,
                                       loglik))))


# Conventional Kriging, Euclidean distances
KC.cst.Ca = krige.control(obj.model = vgmdl.cst.Ca)
kriging.cst.Ca <- krige.conv(obs.gd,
                             data = variable,
                             locations = loc,
                             krige = KC.cst.Ca,
                             dd.dists.mat = ddm$obs,
                             dl.dists.mat = ddm$loc)

```



```{r Ca-cost-based-variogram-plot, fig.width = 6, fig.height = 3, fig.cap = 'Empirical cost-based variogram and fitted model.'}

ggplot(vg.cst.Ca) +
  geom_variogram(vgmdl.cst.Ca) +
  ylab('semivariance Ca')

```


```{r Ca-cost-based-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Cost-based kriging prediction'}

p.Ca <- ggplot(data.frame(loc, Prediction = kriging.cst.Ca$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis()

p.Ca

```



## Comparison of method outcomes

```{r Ca-parameter-table}
kable(par.tab, digits = 2)
```


```{r Ca-compare-variogram-plot, fig.width = 6, fig.height = 4, fig.cap = 'Empirical variogram and fitted models by method for Calcium.'}

vg.both <- rbind(data.frame(vg.std.Ca[1:3],
                            variable = 'Ca',
                            method = 'Euclidean'),
                 data.frame(vg.cst.Ca[1:3],
                            variable = 'Ca',
                            method = 'Cost-based'))

names(vg.both) <- c('distance', 'semivariance', 'n', 'variable', 'method')

p.Ca <- ggplot(filter(vg.both, variable == 'Ca'),
               aes(distance, semivariance)) +
  geom_point(aes(size = n)) + 
  expand_limits(y=0) +
  geom_variogram(vgmdl.std.Ca, variable = 'Ca', method = 'Euclidean') +
  geom_variogram(vgmdl.cst.Ca, variable = 'Ca', method = 'Cost-based') +
  facet_grid(method ~ .) +
  theme_bw()

p.Ca

```




```{r Ca-compare-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Comparison of Kriging estimates.'}

res.df.Ca <- data.frame(loc, 
                        method = c(rep('Classical', nrow(loc)),
                                   rep('Cost-based', nrow(loc)),
                                   rep('Difference', nrow(loc))),
                        Prediction = c(kriging.std.Ca$predict,
                                          kriging.cst.Ca$predict,
                                          kriging.std.Ca$predict-kriging.cst.Ca$predict))

p.Ca <- 
  ggplot(filter(res.df.Ca, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .2, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis() +
  facet_wrap(~method, ncol = 1) 


p.Ca
```


```{r Ca-cost-based-vs-euclidean-prediction, fig.cap = 'Pointwise comparison of predictions by method.'}

## Only represent a random sample of all 30k prediction points
p.Ca <- data.frame(Euclidean  = kriging.std.Ca$predict,
                   Cost_based = kriging.cst.Ca$predict,
                   variable   = 'Ca') %>% 
  sample_n(5e3) %>% 
  ggplot(aes(Euclidean, Cost_based)) +
  geom_point() +
  geom_abline(int = 0, sl = 1, col = 'darkgray') +
  theme_bw()


p.Ca

```

In the scatter plot, the horizontal patterns correspond to predictions on observed values.
Otherwise, the differences are negligeable.

```{r Ca-map-kriging-difference, fig.width=10, fig.height=6, fig.cap='Difference between the Euclidean and the cost-based predictions.'}

pred.dif.Ca <- 
  ggplot(filter(res.df.Ca, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .2, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Difference in\npredictions') +
  scale_color_viridis(name = 'Difference in\npredictions')

pred.dif.Ca
# grid.arrange(pred.comp, pred.dif)
```





```{r Ca-compare-kriging-errors, fig.width=10, fig.height=7, fig.cap = 'Comparison of prediction error by method.'}

res.var.Ca <- data.frame(loc, 
                        method = c(rep('Classical', nrow(loc)),
                                   rep('Cost-based', nrow(loc)),
                                   rep('Difference', nrow(loc))),
                        Prediction = c(kriging.std.Ca$krige.var,
                                          kriging.cst.Ca$krige.var,
                                          kriging.std.Ca$krige.var-kriging.cst.Ca$krige.var))

ggplot(filter(res.var.Ca, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .2, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Kriging\nvariance') +
  scale_color_viridis(name = 'Kriging\nvariance') +
  facet_wrap(~method, ncol = 1) 

```


```{r Ca-map-kriging-difference-errors, fig.width=10, fig.height=6, fig.cap='Difference between the Euclidean and the cost-based prediction errors'}

ggplot(filter(res.var.Ca, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .2, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Difference in\nKriging variance') +
  scale_color_viridis(name = 'Difference in\nKriging variance')
```

Near the observations, the cost-based approach has a larger prediction error due to its increased estimation of the nugget (i.e. short-range variance).
In the main area, the prediction errors are practically the same with both approaches.
Behind the walls, the Euclidean prediction error is unrealistically low.


## Leave-one-out Cross Validation (LOOCV)

```{r Ca-loocv, fig.width = 4, fig.height = 3, fig.cap = 'Pointwise leave-one-out prediction error by method.'}
loocvdat <- obs.gd
loocvdat$data <- variable
res <- sapply(c('std', 'cst'), krige.loocv, loocvdat)
sink()   # Why I need this?

# qplot(res[,1], res[,2]) + geom_abline(int=0, sl=1)

pred_error <- data.frame(
  std = loocvdat$data - res[, 1],
  cst = loocvdat$data - res[, 2]
)

ggplot(pred_error, aes(std, cst)) +
  geom_point() +
  geom_abline(int = 0, sl = 1)

rmse <- function(x) sqrt(mean(x**2))

pred_error %>% 
  gather(method, error) %>% 
  group_by(method) %>% 
  summarise(rmse(error)) %>% 
  kable(digits = 2)

```


\clearpage

# Analysis of Copper



```{r Cu-fix-variable}
variable <- obs.gd$data[, 'Cu']
```


## Euclidean kriging


The variogram model is Exponential.
We choose to estimate the nugget effect, which may account for measurement error, for example.


```{r Cu-euclidean-kriging, include = FALSE}

## compute euclidean (only) variogram
vg.std.Cu <- variog(obs.gd, data = variable)

## fitting variogram models
vgmdl.std.Cu <- likfit(geodata = obs.gd,
                       data = variable,
                       fix.nugget = FALSE,
                       fix.kappa = FALSE,
                       kappa = 0.51,  # If 0.5 then cov.model changes to exponential
                       ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                       cov.model = "exponential",
                       lik.method = "REML")

## Fitted parameters
par.tab <- data.frame(Euclidean = with(vgmdl.std.Cu,
                                       c(beta,
                                         nugget,
                                         sigmasq,
                                         kappa,
                                         phi,
                                         practicalRange)),
                      row.names = c("Intercept",
                                    "Nugget",
                                    "Partial sill",
                                    "kappa",
                                    "phi",
                                    "Pract. range"))


# Conventional Kriging, Euclidean distances
KC.std.Cu = krige.control(obj.model = vgmdl.std.Cu)
kriging.std.Cu <- krige.conv(obs.gd,
                             data = variable,
                             locations = loc,
                             krige = KC.std.Cu)

```

```{r Cu-euclidean-variogram-plot, fig.width = 6,fig.height = 3, fig.cap = 'Empirical variogram and fitted model for Copper.'}

ggplot(vg.std.Cu) +
  geom_variogram(vgmdl.std.Cu) + 
  ylab('semivariance Cu')

```


```{r Cu-euclidean-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Euclidean kriging prediction for Copper.'}

p.Cu <- ggplot(data.frame(loc, Prediction = kriging.std.Cu$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .5, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis()

p.Cu

```



## Cost-based kriging


```{r Cu-cost-based-kriging, include=FALSE}


## compute cost-based empirical variogram
vg.cst.Cu <- variog(obs.gd, data = variable, dists.mat = ddm$obs)

## fitting variogram models
vgmdl.cst.Cu <- likfit(geodata = obs.gd,
                       data = variable,
                       fix.nugget = FALSE,
                       fix.kappa = FALSE,
                       kappa = 0.51,  # If 0.5 then cov.model changes to exponential
                       ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                       cov.model = "exponential",
                       lik.method = "REML",
                       dists.mat = ddm$obs)

## Fitted parameters
par.tab <- cbind(par.tab,
                 data.frame(
                   Cost_based = with(vgmdl.cst.Cu,
                                     c(beta,
                                       nugget,
                                       sigmasq,
                                       kappa,
                                       phi,
                                       practicalRange))))


# Conventional Kriging, Euclidean distances
KC.cst.Cu = krige.control(obj.model = vgmdl.cst.Cu)
kriging.cst.Cu <- krige.conv(obs.gd,
                             data = variable,
                             locations = loc,
                             krige = KC.cst.Cu,
                             dd.dists.mat = ddm$obs,
                             dl.dists.mat = ddm$loc)

```



```{r Cu-cost-based-variogram-plot, fig.width = 6, fig.height = 3, fig.cap = 'Empirical cost-based variogram and fitted model.'}

ggplot(vg.cst.Cu) +
  geom_variogram(vgmdl.cst.Cu) + 
  ylab('semivariance Cu')

```


```{r Cu-cost-based-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Cost-based kriging prediction'}

p.Cu <- ggplot(data.frame(loc, Prediction = kriging.cst.Cu$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .5, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis()

p.Cu

```



## Comparison of method outcomes

```{r Cu-parameter-table}
kable(par.tab, digits = 2)
```


```{r Cu-compare-variogram-plot, fig.width = 6, fig.height = 4, fig.cap = 'Empirical variogram and fitted models by method for Copper.'}

vg.both <- rbind(data.frame(vg.std.Cu[1:3],
                            variable = 'Cu',
                            method = 'Euclidean'),
                 data.frame(vg.cst.Cu[1:3],
                            variable = 'Cu',
                            method = 'Cost-based'))

names(vg.both) <- c('distance', 'semivariance', 'n', 'variable', 'method')

p.Cu <- ggplot(filter(vg.both, variable == 'Cu'),
               aes(distance, semivariance)) +
  geom_point(aes(size = n)) + 
  expand_limits(y=0) +
  geom_variogram(vgmdl.std.Cu, variable = 'Cu', method = 'Euclidean') +
  geom_variogram(vgmdl.cst.Cu, variable = 'Cu', method = 'Cost-based') +
  facet_grid(method ~ .) +
  theme_bw()

p.Cu

```




```{r Cu-compare-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Comparison of Kriging estimates.'}

res.df.Cu <- data.frame(loc, 
                        method = c(rep('Classical', nrow(loc)),
                                   rep('Cost-based', nrow(loc)),
                                   rep('Difference', nrow(loc))),
                        Prediction = c(kriging.std.Cu$predict,
                                       kriging.cst.Cu$predict,
                                       kriging.std.Cu$predict-kriging.cst.Cu$predict))

p.Cu <- 
  ggplot(filter(res.df.Cu, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .5, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis() +
  facet_wrap(~method, ncol = 1) 

p.Cu
```


```{r Cu-cost-based-vs-euclidean-prediction, fig.cap = 'Pointwise comparison of predictions by method.'}

p.Cu <- data.frame(Euclidean  = kriging.std.Cu$predict,
                   Cost_based = kriging.cst.Cu$predict,
                   variable   = 'Cu') %>% 
  sample_n(5e3) %>% 
  ggplot(aes(Euclidean, Cost_based)) +
  geom_point() +
  geom_abline(int = 0, sl = 1, col = 'darkgray') +
  facet_grid(~variable) + 
  theme_bw()

p.Cu

```

In the scatter plot, the horizontal patterns correspond to predictions on observed values.
Otherwise, the differences are negligeable.

```{r Cu-map-kriging-difference, fig.width=10, fig.height=12, fig.cap='Difference between the Euclidean and the cost-based predictions.'}

pred.dif.Cu <- 
  ggplot(filter(res.df.Cu, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .5, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis('Difference in\npredictions') +
  scale_color_viridis('Difference in\npredictions') 

pred.dif.Cu
# grid.arrange(pred.comp, pred.dif)
```




```{r Cu-compare-kriging-errors, fig.width=10, fig.height=7, fig.cap = 'Comparison of prediction error by method.'}

res.var.Cu <- data.frame(loc, 
                        method = c(rep('Classical', nrow(loc)),
                                   rep('Cost-based', nrow(loc)),
                                   rep('Difference', nrow(loc))),
                        Prediction = c(kriging.std.Cu$krige.var,
                                          kriging.cst.Cu$krige.var,
                                          kriging.std.Cu$krige.var-kriging.cst.Cu$krige.var))

ggplot(filter(res.var.Cu, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Kriging\nvariance') +
  scale_color_viridis(name = 'Kriging\nvariance') +
  facet_wrap(~method, ncol = 1) 

```


```{r Cu-map-kriging-difference-errors, fig.width=10, fig.height=6, fig.cap='Difference between the Euclidean and the cost-based prediction errors'}

ggplot(filter(res.var.Cu, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Difference in\nKriging variance') +
  scale_color_viridis(name = 'Difference in\nKriging variance')
```

Near the observations, the cost-based approach has a larger prediction error due to its increased estimation of the nugget (i.e. short-range variance).
In the main area, the prediction errors are practically the same with both approaches.
Behind the walls, the Euclidean prediction error is unrealistically low.




## Leave-one-out Cross Validation (LOOCV)

```{r Cu-loocv, fig.width = 4, fig.height = 3, fig.cap = 'Pointwise leave-one-out prediction error by method.'}
loocvdat <- obs.gd
loocvdat$data <- variable
res <- sapply(c('std', 'cst'), krige.loocv, loocvdat)
sink()   # Why I need this?

# qplot(res[,1], res[,2]) + geom_abline(int=0, sl=1)

pred_error <- data.frame(
  std = loocvdat$data - res[, 1],
  cst = loocvdat$data - res[, 2]
)

ggplot(pred_error, aes(std, cst)) +
  geom_point() +
  geom_abline(int = 0, sl = 1)

rmse <- function(x) sqrt(mean(x**2))

pred_error %>% 
  gather(method, error) %>% 
  group_by(method) %>% 
  summarise(rmse(error)) %>% 
  kable(digits = 2)

```





\clearpage

# Analysis of Ferrum

```{r Fe-fix-variable}
variable <- obs.gd$data[, 'Fe']
```


## Euclidean kriging


The variogram model is Exponential.
We choose to estimate the nugget effect, which may account for measurement error, for example.



```{r Fe-euclidean-kriging, include = FALSE}

## compute euclidean (only) variogram
vg.std.Fe <- variog(obs.gd, data = variable)

## fitting variogram models
vgmdl.std.Fe <- likfit(geodata = obs.gd,
                       data = variable,
                       fix.nugget = FALSE,
                       fix.kappa = FALSE,
                       # kappa = 0.51,  # If 0.5 then cov.model changes to exponential
                       ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                       cov.model = "exponential",
                       lik.method = "REML")

## Fitted parameters
par.tab <- data.frame(
  Euclidean = with(vgmdl.std.Fe,
                   c(beta,
                     nugget,
                     sigmasq,
                     # kappa,
                     phi,
                     practicalRange,
                     loglik)),
  row.names = c("Intercept",
                "Nugget",
                "Partial sill",
                # "kappa",
                "phi",
                "Pract. range",
                "Log-likelihood"))


# Conventional Kriging, Euclidean distances
KC.std.Fe = krige.control(obj.model = vgmdl.std.Fe)
kriging.std.Fe <- krige.conv(obs.gd,
                             data = variable,
                             locations = loc,
                             krige = KC.std.Fe)

```

```{r Fe-euclidean-variogram-plot, fig.width = 6,fig.height = 3, fig.cap = 'Empirical variogram and fitted model.'}

ggplot(vg.std.Fe) + 
  geom_variogram(vgmdl.std.Fe) + 
  ylab('semivariance Fe')

```


```{r Fe-euclidean-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Euclidean kriging prediction'}

p.Fe <- ggplot(data.frame(loc, Prediction = kriging.std.Fe$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis()

p.Fe

```



## Cost-based kriging



```{r Fe-cost-based-kriging, include=FALSE}


## compute cost-based empirical variogram
vg.cst.Fe <- variog(obs.gd, data = variable, dists.mat = ddm$obs)

## fitting variogram models
vgmdl.cst.Fe <- likfit(geodata = obs.gd,
                       data = variable,
                       fix.nugget = FALSE,
                       fix.kappa = FALSE,
                       kappa = 0.51,  # If 0.5 then cov.model changes to exponential
                       ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                       cov.model = "exponential",
                       lik.method = "REML",
                       dists.mat = ddm$obs)

## Fitted parameters
par.tab <- cbind(par.tab,
                 data.frame(
                   Cost_based = with(vgmdl.cst.Fe,
                                     c(beta,
                                       nugget,
                                       sigmasq,
                                       # kappa,
                                       phi,
                                       practicalRange,
                                       loglik))))


# Conventional Kriging, Euclidean distances
KC.cst.Fe = krige.control(obj.model = vgmdl.cst.Fe)
kriging.cst.Fe <- krige.conv(obs.gd,
                             data = variable,
                             locations = loc,
                             krige = KC.cst.Fe,
                             dd.dists.mat = ddm$obs,
                             dl.dists.mat = ddm$loc)

```



```{r Fe-cost-based-variogram-plot, fig.width = 6, fig.height = 3, fig.cap = 'Empirical cost-based variogram and fitted model.'}

ggplot(vg.cst.Fe) +
  geom_variogram(vgmdl.cst.Fe) +
  ylab('semivariance Fe')

```


```{r Fe-cost-based-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Cost-based kriging prediction'}

p.Fe <- ggplot(data.frame(loc, Prediction = kriging.cst.Fe$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis()

p.Fe

```



## Comparison of method outcomes

```{r Fe-parameter-table}
kable(par.tab, digits = 2)
```


```{r Fe-compare-variogram-plot, fig.width = 6, fig.height = 4, fig.cap = 'Empirical variogram and fitted models by method for Ferrum.'}

vg.both <- rbind(data.frame(vg.std.Fe[1:3],
                            variable = 'Fe',
                            method = 'Euclidean'),
                 data.frame(vg.cst.Fe[1:3],
                            variable = 'Fe',
                            method = 'Cost-based'))

names(vg.both) <- c('distance', 'semivariance', 'n', 'variable', 'method')

p.Fe <- ggplot(filter(vg.both, variable == 'Fe'),
               aes(distance, semivariance)) +
  geom_point(aes(size = n)) + 
  expand_limits(y=0) +
  geom_variogram(vgmdl.std.Fe, variable = 'Fe', method = 'Euclidean') +
  geom_variogram(vgmdl.cst.Fe, variable = 'Fe', method = 'Cost-based') +
  facet_grid(method ~ .) +
  theme_bw()

p.Fe

```




```{r Fe-compare-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Comparison of Kriging estimates.'}

res.df.Fe <- data.frame(loc, 
                        method = c(rep('Classical', nrow(loc)),
                                   rep('Cost-based', nrow(loc)),
                                   rep('Difference', nrow(loc))),
                        Prediction = c(kriging.std.Fe$predict,
                                          kriging.cst.Fe$predict,
                                          kriging.std.Fe$predict-kriging.cst.Fe$predict))

p.Fe <- 
  ggplot(filter(res.df.Fe, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .05, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis() +
  facet_wrap(~method, ncol = 1) 


p.Fe
```


```{r Fe-cost-based-vs-euclidean-prediction, fig.cap = 'Pointwise comparison of predictions by method.'}

p.Fe <- data.frame(Euclidean  = kriging.std.Fe$predict,
                   Cost_based = kriging.cst.Fe$predict) %>% 
  sample_n(5e3) %>% 
  ggplot(aes(Euclidean, Cost_based)) +
  geom_point() +
  geom_abline(int = 0, sl = 1, col = 'darkgray') +
  theme_bw()


p.Fe

```

In the scatter plot, the horizontal patterns correspond to predictions on observed values.
Otherwise, the differences are negligeable.

```{r Fe-map-kriging-difference, fig.width=10, fig.height=6, fig.cap='Difference between the Euclidean and the cost-based predictions.'}

pred.dif.Fe <- 
  ggplot(filter(res.df.Fe, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .05, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Difference in\npredictions') +
  scale_color_viridis(name = 'Difference in\npredictions')

pred.dif.Fe
# grid.arrange(pred.comp, pred.dif)
```





```{r Fe-compare-kriging-errors, fig.width=10, fig.height=7, fig.cap = 'Comparison of prediction error by method.'}

res.var.Fe <- data.frame(loc, 
                        method = c(rep('Classical', nrow(loc)),
                                   rep('Cost-based', nrow(loc)),
                                   rep('Difference', nrow(loc))),
                        Prediction = c(kriging.std.Fe$krige.var,
                                          kriging.cst.Fe$krige.var,
                                          kriging.std.Fe$krige.var-kriging.cst.Fe$krige.var))

ggplot(filter(res.var.Fe, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .05, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Kriging\nvariance') +
  scale_color_viridis(name = 'Kriging\nvariance') +
  facet_wrap(~method, ncol = 1) 

```


```{r Fe-map-kriging-difference-errors, fig.width=10, fig.height=6, fig.cap='Difference between the Euclidean and the cost-based prediction errors'}

ggplot(filter(res.var.Fe, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .05, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Difference in\nKriging variance') +
  scale_color_viridis(name = 'Difference in\nKriging variance')
```

Near the observations, the cost-based approach has a larger prediction error due to its increased estimation of the nugget (i.e. short-range variance).
In the main area, the prediction errors are practically the same with both approaches.
Behind the walls, the Euclidean prediction error is unrealistically low.


## Leave-one-out Cross Validation (LOOCV)

```{r Fe-loocv, fig.width = 4, fig.height = 3, fig.cap = 'Pointwise leave-one-out prediction error by method.'}
loocvdat <- obs.gd
loocvdat$data <- variable
res <- sapply(c('std', 'cst'), krige.loocv, loocvdat)
sink()   # Why I need this?

# qplot(res[,1], res[,2]) + geom_abline(int=0, sl=1)

pred_error <- data.frame(
  std = loocvdat$data - res[, 1],
  cst = loocvdat$data - res[, 2]
)

ggplot(pred_error, aes(std, cst)) +
  geom_point() +
  geom_abline(int = 0, sl = 1)

rmse <- function(x) sqrt(mean(x**2))

pred_error %>% 
  gather(method, error) %>% 
  group_by(method) %>% 
  summarise(rmse(error)) %>% 
  kable(digits = 2)

```






\clearpage

# Analysis of Potasium

```{r K-fix-variable}
variable <- obs.gd$data[, 'K']
```


## Euclidean kriging


The variogram model is Exponential.
We choose to estimate the nugget effect, which may account for measurement error, for example.



```{r K-euclidean-kriging, include = FALSE}

## compute euclidean (only) variogram
vg.std <- variog(obs.gd, data = variable)

## fitting variogram models
vgmdl.std <- likfit(geodata = obs.gd,
                       data = variable,
                       fix.nugget = FALSE,
                       fix.kappa = FALSE,
                       # kappa = 0.51,  # If 0.5 then cov.model changes to exponential
                       ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                       cov.model = "exponential",
                       lik.method = "REML")

## Fitted parameters
par.tab <- data.frame(
  Euclidean = with(vgmdl.std,
                   c(beta,
                     nugget,
                     sigmasq,
                     # kappa,
                     phi,
                     practicalRange,
                     loglik)),
  row.names = c("Intercept",
                "Nugget",
                "Partial sill",
                # "kappa",
                "phi",
                "Pract. range",
                "Log-likelihood"))


# Conventional Kriging, Euclidean distances
KC.std = krige.control(obj.model = vgmdl.std)
kriging.std <- krige.conv(obs.gd,
                             data = variable,
                             locations = loc,
                             krige = KC.std)

```

```{r K-euclidean-variogram-plot, fig.width = 6,fig.height = 3, fig.cap = 'Empirical variogram and fitted model.'}

ggplot(vg.std) + 
  geom_variogram(vgmdl.std) + 
  ylab('semivariance K')

```


```{r K-euclidean-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Euclidean kriging prediction'}

p <- ggplot(data.frame(loc, Prediction = kriging.std$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis()

p

```



## Cost-based kriging



```{r K-cost-based-kriging, include=FALSE}


## compute cost-based empirical variogram
vg.cst <- variog(obs.gd, data = variable, dists.mat = ddm$obs)

## fitting variogram models
vgmdl.cst <- likfit(geodata = obs.gd,
                       data = variable,
                       fix.nugget = FALSE,
                       fix.kappa = FALSE,
                       kappa = 0.51,  # If 0.5 then cov.model changes to exponential
                       ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                       cov.model = "exponential",
                       lik.method = "REML",
                       dists.mat = ddm$obs)

## Fitted parameters
par.tab <- cbind(par.tab,
                 data.frame(
                   Cost_based = with(vgmdl.cst,
                                     c(beta,
                                       nugget,
                                       sigmasq,
                                       # kappa,
                                       phi,
                                       practicalRange,
                                       loglik))))


# Conventional Kriging, Euclidean distances
KC.cst = krige.control(obj.model = vgmdl.cst)
kriging.cst <- krige.conv(obs.gd,
                             data = variable,
                             locations = loc,
                             krige = KC.cst,
                             dd.dists.mat = ddm$obs,
                             dl.dists.mat = ddm$loc)

```



```{r K-cost-based-variogram-plot, fig.width = 6, fig.height = 3, fig.cap = 'Empirical cost-based variogram and fitted model.'}

ggplot(vg.cst) +
  geom_variogram(vgmdl.cst) +
  ylab('semivariance K')

```


```{r K-cost-based-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Cost-based kriging prediction'}

p <- ggplot(data.frame(loc, Prediction = kriging.cst$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis()

p

```



## Comparison of method outcomes

```{r K-parameter-table}
kable(par.tab, digits = 2)
```


```{r K-compare-variogram-plot, fig.width = 6, fig.height = 4, fig.cap = 'Empirical variogram and fitted models by method for Potasium.'}

vg.both <- rbind(data.frame(vg.std[1:3],
                            variable = 'K',
                            method = 'Euclidean'),
                 data.frame(vg.cst[1:3],
                            variable = 'K',
                            method = 'Cost-based'))

names(vg.both) <- c('distance', 'semivariance', 'n', 'variable', 'method')

p <- ggplot(filter(vg.both, variable == 'K'),
               aes(distance, semivariance)) +
  geom_point(aes(size = n)) + 
  expand_limits(y=0) +
  geom_variogram(vgmdl.std, variable = 'K', method = 'Euclidean') +
  geom_variogram(vgmdl.cst, variable = 'K', method = 'Cost-based') +
  facet_grid(method ~ .) +
  theme_bw()

p

```




```{r K-compare-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Comparison of Kriging estimates.'}

res.df <- data.frame(loc, 
                        method = c(rep('Classical', nrow(loc)),
                                   rep('Cost-based', nrow(loc)),
                                   rep('Difference', nrow(loc))),
                        Prediction = c(kriging.std$predict,
                                          kriging.cst$predict,
                                          kriging.std$predict-kriging.cst$predict))

p <- 
  ggplot(filter(res.df, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .01, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis() +
  facet_wrap(~method, ncol = 1) 


p
```


```{r K-cost-based-vs-euclidean-prediction, fig.cap = 'Pointwise comparison of predictions by method.'}

p <- data.frame(Euclidean  = kriging.std$predict,
                   Cost_based = kriging.cst$predict) %>% 
  sample_n(5e3) %>% 
  ggplot(aes(Euclidean, Cost_based)) +
  geom_point() +
  geom_abline(int = 0, sl = 1, col = 'darkgray') +
  theme_bw()


p

```

In the scatter plot, the horizontal patterns correspond to predictions on observed values.
Otherwise, the differences are negligeable.

```{r K-map-kriging-difference, fig.width=10, fig.height=6, fig.cap='Difference between the Euclidean and the cost-based predictions.'}

pred.dif <- 
  ggplot(filter(res.df, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .01, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Difference in\npredictions') +
  scale_color_viridis(name = 'Difference in\npredictions')

pred.dif
# grid.arrange(pred.comp, pred.dif)
```





```{r K-compare-kriging-errors, fig.width=10, fig.height=7, fig.cap = 'Comparison of prediction error by method.'}

res.var <- data.frame(loc, 
                        method = c(rep('Classical', nrow(loc)),
                                   rep('Cost-based', nrow(loc)),
                                   rep('Difference', nrow(loc))),
                        Prediction = c(kriging.std$krige.var,
                                          kriging.cst$krige.var,
                                          kriging.std$krige.var-kriging.cst$krige.var))

ggplot(filter(res.var, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .01, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Kriging\nvariance') +
  scale_color_viridis(name = 'Kriging\nvariance') +
  facet_wrap(~method, ncol = 1) 

```


```{r K-map-kriging-difference-errors, fig.width=10, fig.height=6, fig.cap='Difference between the Euclidean and the cost-based prediction errors'}

ggplot(filter(res.var, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .01, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Difference in\nKriging variance') +
  scale_color_viridis(name = 'Difference in\nKriging variance')
```

Near the observations, the cost-based approach has a larger prediction error due to its increased estimation of the nugget (i.e. short-range variance).
In the main area, the prediction errors are practically the same with both approaches.
Behind the walls, the Euclidean prediction error is unrealistically low.


## Leave-one-out Cross Validation (LOOCV)

```{r K-loocv, fig.width = 4, fig.height = 3, fig.cap = 'Pointwise leave-one-out prediction error by method.'}
loocvdat <- obs.gd
loocvdat$data <- variable
res <- sapply(c('std', 'cst'), krige.loocv, loocvdat)
sink()   # Why I need this?

# qplot(res[,1], res[,2]) + geom_abline(int=0, sl=1)

pred_error <- data.frame(
  std = loocvdat$data - res[, 1],
  cst = loocvdat$data - res[, 2]
)

ggplot(pred_error, aes(std, cst)) +
  geom_point() +
  geom_abline(int = 0, sl = 1)

rmse <- function(x) sqrt(mean(x**2))

pred_error %>% 
  gather(method, error) %>% 
  group_by(method) %>% 
  summarise(rmse(error)) %>% 
  kable(digits = 2)

```







\clearpage

# Analysis of Magnesium

```{r Mg-fix-variable}
variable <- obs.gd$data[, 'Mg']
```


## Euclidean kriging


The variogram model is Exponential.
We choose to estimate the nugget effect, which may account for measurement error, for example.



```{r Mg-euclidean-kriging, include = FALSE}

## compute euclidean (only) variogram
vg.std.Mg <- variog(obs.gd, data = variable)

## fitting variogram models
vgmdl.std.Mg <- likfit(geodata = obs.gd,
                       data = variable,
                       fix.nugget = FALSE,
                       fix.kappa = FALSE,
                       # kappa = 0.51,  # If 0.5 then cov.model changes to exponential
                       ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                       cov.model = "exponential",
                       lik.method = "REML")

## Fitted parameters
par.tab <- data.frame(
  Euclidean = with(vgmdl.std.Mg,
                   c(beta,
                     nugget,
                     sigmasq,
                     # kappa,
                     phi,
                     practicalRange,
                     loglik)),
  row.names = c("Intercept",
                "Nugget",
                "Partial sill",
                # "kappa",
                "phi",
                "Pract. range",
                "Log-likelihood"))


# Conventional Kriging, Euclidean distances
KC.std.Mg = krige.control(obj.model = vgmdl.std.Mg)
kriging.std.Mg <- krige.conv(obs.gd,
                             data = variable,
                             locations = loc,
                             krige = KC.std.Mg)

```

```{r Mg-euclidean-variogram-plot, fig.width = 6,fig.height = 3, fig.cap = 'Empirical variogram and fitted model.'}

ggplot(vg.std.Mg) + 
  geom_variogram(vgmdl.std.Mg) + 
  ylab('semivariance Mg')

```


```{r Mg-euclidean-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Euclidean kriging prediction'}

p.Mg <- ggplot(data.frame(loc, Prediction = kriging.std.Mg$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis()

p.Mg

```



## Cost-based kriging



```{r Mg-cost-based-kriging, include=FALSE}


## compute cost-based empirical variogram
vg.cst.Mg <- variog(obs.gd, data = variable, dists.mat = ddm$obs)

## fitting variogram models
vgmdl.cst.Mg <- likfit(geodata = obs.gd,
                       data = variable,
                       fix.nugget = FALSE,
                       fix.kappa = FALSE,
                       kappa = 0.51,  # If 0.5 then cov.model changes to exponential
                       ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                       cov.model = "exponential",
                       lik.method = "REML",
                       dists.mat = ddm$obs)

## Fitted parameters
par.tab <- cbind(par.tab,
                 data.frame(
                   Cost_based = with(vgmdl.cst.Mg,
                                     c(beta,
                                       nugget,
                                       sigmasq,
                                       # kappa,
                                       phi,
                                       practicalRange,
                                       loglik))))


# Conventional Kriging, Euclidean distances
KC.cst.Mg = krige.control(obj.model = vgmdl.cst.Mg)
kriging.cst.Mg <- krige.conv(obs.gd,
                             data = variable,
                             locations = loc,
                             krige = KC.cst.Mg,
                             dd.dists.mat = ddm$obs,
                             dl.dists.mat = ddm$loc)

```



```{r Mg-cost-based-variogram-plot, fig.width = 6, fig.height = 3, fig.cap = 'Empirical cost-based variogram and fitted model.'}

ggplot(vg.cst.Mg) +
  geom_variogram(vgmdl.cst.Mg) +
  ylab('semivariance Mg')

```


```{r Mg-cost-based-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Cost-based kriging prediction'}

p.Mg <- ggplot(data.frame(loc, Prediction = kriging.cst.Mg$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis()

p.Mg

```



## Comparison of method outcomes

```{r Mg-parameter-table}
kable(par.tab, digits = 2)
```


```{r Mg-compare-variogram-plot, fig.width = 6, fig.height = 4, fig.cap = 'Empirical variogram and fitted models by method for Magnesium.'}

vg.both <- rbind(data.frame(vg.std.Mg[1:3],
                            variable = 'Mg',
                            method = 'Euclidean'),
                 data.frame(vg.cst.Mg[1:3],
                            variable = 'Mg',
                            method = 'Cost-based'))

names(vg.both) <- c('distance', 'semivariance', 'n', 'variable', 'method')

p.Mg <- ggplot(filter(vg.both, variable == 'Mg'),
               aes(distance, semivariance)) +
  geom_point(aes(size = n)) + 
  expand_limits(y=0) +
  geom_variogram(vgmdl.std.Mg, variable = 'Mg', method = 'Euclidean') +
  geom_variogram(vgmdl.cst.Mg, variable = 'Mg', method = 'Cost-based') +
  facet_grid(method ~ .) +
  theme_bw()

p.Mg

```




```{r Mg-compare-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Comparison of Kriging estimates.'}

res.df.Mg <- data.frame(loc, 
                        method = c(rep('Classical', nrow(loc)),
                                   rep('Cost-based', nrow(loc)),
                                   rep('Difference', nrow(loc))),
                        Prediction = c(kriging.std.Mg$predict,
                                          kriging.cst.Mg$predict,
                                          kriging.std.Mg$predict-kriging.cst.Mg$predict))

p.Mg <- 
  ggplot(filter(res.df.Mg, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .02, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis() +
  facet_wrap(~method, ncol = 1) 


p.Mg
```


```{r Mg-cost-based-vs-euclidean-prediction, fig.cap = 'Pointwise comparison of predictions by method.'}

p.Mg <- data.frame(Euclidean  = kriging.std.Mg$predict,
                   Cost_based = kriging.cst.Mg$predict) %>% 
  sample_n(5e3) %>% 
  ggplot(aes(Euclidean, Cost_based)) +
  geom_point() +
  geom_abline(int = 0, sl = 1, col = 'darkgray') +
  theme_bw()


p.Mg

```

In the scatter plot, the horizontal patterns correspond to predictions on observed values.
Otherwise, the differences are negligeable.

```{r Mg-map-kriging-difference, fig.width=10, fig.height=6, fig.cap='Difference between the Euclidean and the cost-based predictions.'}

pred.dif.Mg <- 
  ggplot(filter(res.df.Mg, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .02, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Difference in\npredictions') +
  scale_color_viridis(name = 'Difference in\npredictions')

pred.dif.Mg
# grid.arrange(pred.comp, pred.dif)
```





```{r Mg-compare-kriging-errors, fig.width=10, fig.height=7, fig.cap = 'Comparison of prediction error by method.'}

res.var.Mg <- data.frame(loc, 
                        method = c(rep('Classical', nrow(loc)),
                                   rep('Cost-based', nrow(loc)),
                                   rep('Difference', nrow(loc))),
                        Prediction = c(kriging.std.Mg$krige.var,
                                          kriging.cst.Mg$krige.var,
                                          kriging.std.Mg$krige.var-kriging.cst.Mg$krige.var))

ggplot(filter(res.var.Mg, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .02, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Kriging\nvariance') +
  scale_color_viridis(name = 'Kriging\nvariance') +
  facet_wrap(~method, ncol = 1) 

```


```{r Mg-map-kriging-difference-errors, fig.width=10, fig.height=6, fig.cap='Difference between the Euclidean and the cost-based prediction errors'}

ggplot(filter(res.var.Mg, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .02, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Difference in\nKriging variance') +
  scale_color_viridis(name = 'Difference in\nKriging variance')
```

Near the observations, the cost-based approach has a larger prediction error due to its increased estimation of the nugget (i.e. short-range variance).
In the main area, the prediction errors are practically the same with both approaches.
Behind the walls, the Euclidean prediction error is unrealistically low.


## Leave-one-out Cross Validation (LOOCV)

```{r Mg-loocv, fig.width = 4, fig.height = 3, fig.cap = 'Pointwise leave-one-out prediction error by method.'}
loocvdat <- obs.gd
loocvdat$data <- variable
res <- sapply(c('std', 'cst'), krige.loocv, loocvdat)
sink()   # Why I need this?

# qplot(res[,1], res[,2]) + geom_abline(int=0, sl=1)

pred_error <- data.frame(
  std = loocvdat$data - res[, 1],
  cst = loocvdat$data - res[, 2]
)

ggplot(pred_error, aes(std, cst)) +
  geom_point() +
  geom_abline(int = 0, sl = 1)

rmse <- function(x) sqrt(mean(x**2))

pred_error %>% 
  gather(method, error) %>% 
  group_by(method) %>% 
  summarise(rmse(error)) %>% 
  kable(digits = 2)

```








\clearpage

# Analysis of Zinc

```{r Zn-fix-variable}
variable <- obs.gd$data[, 'Zn']
```


## Euclidean kriging


The variogram model is Exponential.
We choose to estimate the nugget effect, which may account for measurement error, for example.



```{r Zn-euclidean-kriging, include = FALSE}

## compute euclidean (only) variogram
vg.std.Zn <- variog(obs.gd, data = variable)

## fitting variogram models
vgmdl.std.Zn <- likfit(geodata = obs.gd,
                       data = variable,
                       fix.nugget = FALSE,
                       fix.kappa = FALSE,
                       # kappa = 0.51,  # If 0.5 then cov.model changes to exponential
                       ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                       cov.model = "exponential",
                       lik.method = "REML")

## Fitted parameters
par.tab <- data.frame(
  Euclidean = with(vgmdl.std.Zn,
                   c(beta,
                     nugget,
                     sigmasq,
                     # kappa,
                     phi,
                     practicalRange,
                     loglik)),
  row.names = c("Intercept",
                "Nugget",
                "Partial sill",
                # "kappa",
                "phi",
                "Pract. range",
                "Log-likelihood"))


# Conventional Kriging, Euclidean distances
KC.std.Zn = krige.control(obj.model = vgmdl.std.Zn)
kriging.std.Zn <- krige.conv(obs.gd,
                             data = variable,
                             locations = loc,
                             krige = KC.std.Zn)

```

```{r Zn-euclidean-variogram-plot, fig.width = 6,fig.height = 3, fig.cap = 'Empirical variogram and fitted model.'}

ggplot(vg.std.Zn) + 
  geom_variogram(vgmdl.std.Zn) + 
  ylab('semivariance Zn')

```


```{r Zn-euclidean-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Euclidean kriging prediction'}

p.Zn <- ggplot(data.frame(loc, Prediction = kriging.std.Zn$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis()

p.Zn

```



## Cost-based kriging



```{r Zn-cost-based-kriging, include=FALSE}


## compute cost-based empirical variogram
vg.cst.Zn <- variog(obs.gd, data = variable, dists.mat = ddm$obs)

## fitting variogram models
vgmdl.cst.Zn <- likfit(geodata = obs.gd,
                       data = variable,
                       fix.nugget = FALSE,
                       fix.kappa = FALSE,
                       kappa = 0.51,  # If 0.5 then cov.model changes to exponential
                       ini = c(10, 5),  # sigma^2 (partial sill) and phi (range parameter)
                       cov.model = "exponential",
                       lik.method = "REML",
                       dists.mat = ddm$obs)

## Fitted parameters
par.tab <- cbind(par.tab,
                 data.frame(
                   Cost_based = with(vgmdl.cst.Zn,
                                     c(beta,
                                       nugget,
                                       sigmasq,
                                       # kappa,
                                       phi,
                                       practicalRange,
                                       loglik))))


# Conventional Kriging, Euclidean distances
KC.cst.Zn = krige.control(obj.model = vgmdl.cst.Zn)
kriging.cst.Zn <- krige.conv(obs.gd,
                             data = variable,
                             locations = loc,
                             krige = KC.cst.Zn,
                             dd.dists.mat = ddm$obs,
                             dl.dists.mat = ddm$loc)

```



```{r Zn-cost-based-variogram-plot, fig.width = 6, fig.height = 3, fig.cap = 'Empirical cost-based variogram and fitted model.'}

ggplot(vg.cst.Zn) +
  geom_variogram(vgmdl.cst.Zn) +
  ylab('semivariance Zn')

```


```{r Zn-cost-based-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Cost-based kriging prediction'}

p.Zn <- ggplot(data.frame(loc, Prediction = kriging.cst.Zn$predict), aes(x, y)) +
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis()

p.Zn

```



## Comparison of method outcomes

```{r Zn-parameter-table}
kable(par.tab, digits = 2)
```


```{r Zn-compare-variogram-plot, fig.width = 6, fig.height = 4, fig.cap = 'Empirical variogram and fitted models by method for Zinc.'}

vg.both <- rbind(data.frame(vg.std.Zn[1:3],
                            variable = 'Zn',
                            method = 'Euclidean'),
                 data.frame(vg.cst.Zn[1:3],
                            variable = 'Zn',
                            method = 'Cost-based'))

names(vg.both) <- c('distance', 'semivariance', 'n', 'variable', 'method')

p.Zn <- ggplot(filter(vg.both, variable == 'Zn'),
               aes(distance, semivariance)) +
  geom_point(aes(size = n)) + 
  expand_limits(y=0) +
  geom_variogram(vgmdl.std.Zn, variable = 'Zn', method = 'Euclidean') +
  geom_variogram(vgmdl.cst.Zn, variable = 'Zn', method = 'Cost-based') +
  facet_grid(method ~ .) +
  theme_bw()

p.Zn

```




```{r Zn-compare-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Comparison of Kriging estimates.'}

res.df.Zn <- data.frame(loc, 
                        method = c(rep('Classical', nrow(loc)),
                                   rep('Cost-based', nrow(loc)),
                                   rep('Difference', nrow(loc))),
                        Prediction = c(kriging.std.Zn$predict,
                                          kriging.cst.Zn$predict,
                                          kriging.std.Zn$predict-kriging.cst.Zn$predict))

p.Zn <- 
  ggplot(filter(res.df.Zn, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .5, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis() +
  scale_color_viridis() +
  facet_wrap(~method, ncol = 1) 


p.Zn
```


```{r Zn-cost-based-vs-euclidean-prediction, fig.cap = 'Pointwise comparison of predictions by method.'}

p.Zn <- data.frame(Euclidean  = kriging.std.Zn$predict,
                   Cost_based = kriging.cst.Zn$predict) %>% 
  sample_n(5e3) %>% 
  ggplot(aes(Euclidean, Cost_based)) +
  geom_point() +
  geom_abline(int = 0, sl = 1, col = 'darkgray') +
  theme_bw()


p.Zn

```

In the scatter plot, the horizontal patterns correspond to predictions on observed values.
Otherwise, the differences are negligeable.

```{r Zn-map-kriging-difference, fig.width=10, fig.height=6, fig.cap='Difference between the Euclidean and the cost-based predictions.'}

pred.dif.Zn <- 
  ggplot(filter(res.df.Zn, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .5, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Difference in\npredictions') +
  scale_color_viridis(name = 'Difference in\npredictions')

pred.dif.Zn
# grid.arrange(pred.comp, pred.dif)
```





```{r Zn-compare-kriging-errors, fig.width=10, fig.height=7, fig.cap = 'Comparison of prediction error by method.'}

res.var.Zn <- data.frame(loc, 
                        method = c(rep('Classical', nrow(loc)),
                                   rep('Cost-based', nrow(loc)),
                                   rep('Difference', nrow(loc))),
                        Prediction = c(kriging.std.Zn$krige.var,
                                          kriging.cst.Zn$krige.var,
                                          kriging.std.Zn$krige.var-kriging.cst.Zn$krige.var))

ggplot(filter(res.var.Zn, method != 'Difference'),
         aes(x, y)) + 
  geom_raster(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .5, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Kriging\nvariance') +
  scale_color_viridis(name = 'Kriging\nvariance') +
  facet_wrap(~method, ncol = 1) 

```


```{r Zn-map-kriging-difference-errors, fig.width=10, fig.height=6, fig.cap='Difference between the Euclidean and the cost-based prediction errors'}

ggplot(filter(res.var.Zn, method == 'Difference'),
         aes(x, y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), binwidth = .5, color = 'lightgray') +
  geom_india() +
  scale_fill_viridis(name = 'Difference in\nKriging variance') +
  scale_color_viridis(name = 'Difference in\nKriging variance')
```

Near the observations, the cost-based approach has a larger prediction error due to its increased estimation of the nugget (i.e. short-range variance).
In the main area, the prediction errors are practically the same with both approaches.
Behind the walls, the Euclidean prediction error is unrealistically low.


## Leave-one-out Cross Validation (LOOCV)

```{r Zn-loocv, fig.width = 4, fig.height = 3, fig.cap = 'Pointwise leave-one-out prediction error by method.'}
loocvdat <- obs.gd
loocvdat$data <- variable
res <- sapply(c('std', 'cst'), krige.loocv, loocvdat)
sink()   # Why I need this?

# qplot(res[,1], res[,2]) + geom_abline(int=0, sl=1)

pred_error <- data.frame(
  std = loocvdat$data - res[, 1],
  cst = loocvdat$data - res[, 2]
)

ggplot(pred_error, aes(std, cst)) +
  geom_point() +
  geom_abline(int = 0, sl = 1)

rmse <- function(x) sqrt(mean(x**2))

pred_error %>% 
  gather(method, error) %>% 
  group_by(method) %>% 
  summarise(rmse(error)) %>% 
  kable(digits = 2)

```






\clearpage

# Conclusions

- Actually, the kriging model for Calcium is not adjusting very well the tails of the data, which are heavier than expected.
  This happens both for the Euclidean and cost-based models. This means that none of both approaches will be really good predictors anyways.

```{r model-diagnostics, fig.width = 6, fig.height = 6, fig.cap = 'Diagnostics of residuals for Calcium.'}
## Not really a nice fit
resid.Ca <- rbind(data.frame(x = resid(vgmdl.std.Ca),
                             model = 'Euclidean'),
                  data.frame(x = resid(vgmdl.cst.Ca),
                             model = 'Cost-based'))

## manual qqline computation
qql <- function(x, probs = c(0.25, 0.75)) {
  sq <- quantile(x, probs)  # sample quantiles
  tq <- qnorm(probs)                 # theoretical quantiles
  slp <- diff(sq)/diff(tq)
  data.frame(intercept = sq[1L] - slp * tq[1L],
             slope = slp)
}

resid.Ca %>% 
  group_by(model) %>% 
  summarize(int = qql(x)$intercept,
            slp = qql(x)$slope) -> dat.lines

grid.arrange(
  ggplot(resid.Ca, aes(x)) +
    geom_histogram() +
    xlab('Residuals Ca') + ylab(NULL) +
    facet_wrap(~ model),
  ggplot(resid.Ca, aes(sample = x)) +
    geom_point(stat = 'qq') +
    geom_abline(aes(int = int, slope = slp), data = dat.lines) +
    facet_wrap(~ model)
)

```

- The rest of the variables do not show a clear discontinuity as a consequence of the walls. The Euclidean and cost-based predictions are very similar. Furthermore, the Euclidean and cost-based emprirical variograms display practically the same shape in all cases. This suggests that the solid structures are not really affecting the spatial distribution of chemicals.

- Many variables display an initial drop in the semivariance. Even this is based on only 10 or 12 pairs of observations in the first lag, the drop is dramatic for several variables like Copper, Potasium or Magnesium. This may be due to the presence of some extreme values which contrast heavily with neighbouring values. The impact in higher lags is absorbed by the high number of pairs.
