---
title: "Comparison of classical kriging vs cost-based kriging"
author: "Facundo Muñoz"
date: "26/03/2015"
output:
  pdf_document:
    fig_caption: true
    pandoc_args: [
      "--output", "../reports/Kriging_comparison.pdf"
    ]
---

```{r ProjectTemplate-setup, include = FALSE}
## This automatically loads packages, data and helper functions
## and perform munging operations as defined for the project
library(ProjectTemplate)
load.project()
```

Here we compare the outcome of a classical kriging against a cost-based kriging which takes into account the presence of a semi-barrier

## Data description

```{r setup-knitr, include = FALSE}
opts_chunk$set(echo       = FALSE,
               message    = FALSE,
               warning    = FALSE,
               comment    = NA,
               fig.width  = 6,
               fig.height = 5,
               fig.caption = TRUE,
               cache      = TRUE)
```

Read coordinates from csv.
Units are meters? Rescale to km.

```{r load, echo=FALSE}
path <- "/home/facu/Work/Proyectos/JNegre-Arqueology"
fauna_coord <- read.csv(file.path(path, 'data', 'faunacoordinates.csv'),
                        header = FALSE,
                        col.names = c('x', 'y'))
fauna_coord <- fauna_coord/1e3

raw_dat <- read_excel(file.path(path, 'data', 'Multielements coordinates.xlsx'),
                      skip = 1)

multielem <- raw_dat[,  1:13] %>%
  filter(!is.na(X)) %>%
  mutate(X = X/1e3, Y = Y/1e3) %>%
  as.data.frame

## Little innocent manipulation
multielem[1, 'Ca %'] <- 3

matorg <- raw_dat[,  -(1:13)] %>%
  filter(!is.na(X)) %>%
  mutate(X = X/1e3, Y = Y/1e3)
```


I modified the value of Calcium for observation 1, to improve the effect of the visualization



```{r functions, echo=FALSE}

## Distance from one point to a whole set of points
dist1 <- function(o, x) {
  x <- as.matrix(x)
  n <- nrow(x)
  sqrt(apply((x - matrix(1, nrow = n) %*% o)**2, 1, sum))
}


```

Take a look at the data and build a fake circular border at the bottom-left corner.

I left small gap (say, a door) in order to emphasize the difference between classical and cost-based estimations.

```{r plot-coord, fig.cap = 'Observation locations and strong barrier. In gray, the locations of the bones.'}
coordinates(multielem) <- ~ X + Y

## Build a regular grid spanning the region
# Define region as a rectangle containing all the points
# with limits extended by 4%
extend.range <- function(x, pct) {
  range <- range(x)
  range + c(-1, 1) * pct * diff(range) / 100
}
pct = 4
lims <- sapply(data.frame(coordinates(multielem)), extend.range, pct)

## Definition of the hut
origin <- lims[1, ]
radius <- dist1(origin, coordinates(multielem)) %>%
  min %>%
  prod(.95) %>%
  round(1)

## Trace a circular border with the given origin and radius
border_resolution <- 51
theta_seq <- seq(0, pi/2 - pi/18, length = border_resolution)
border_coord <- cbind(x = cos(theta_seq), y = sin(theta_seq))
border_coord <- matrix(1, nrow=2*border_resolution+1) %*% origin +
  rbind(radius*1.025*border_coord[1,],
        radius*.975*border_coord,
        radius*1.025*border_coord[border_resolution:1,]) %>%
  data.frame()


ggplot(fauna_coord, aes(x, y)) +
  geom_point(col = 'lightgray') +
  geom_point(aes(X, Y), data = data.frame(multielem)) +
  geom_polygon(aes(x, y), data = border_coord) +
  coord_fixed() +
  theme_bw()
```


## Variogram estimation

Empirical (classical) variogram with Calcium observations.
No need to compute cost-based variogram, as the only obstacle is off the observations envelope.

The variogram model is exponential with nugget fixed to 0, as is expected for a continuously varying quantity like Calcium content.
In any case, it is estimated as zero.

```{r variogram, include=FALSE}
## Observations as geodata
obs.gd <- as.geodata(multielem, data.col="Ca %")

## Distance matrix Obs-Obs
## No obstacles within the observations
## So, no need to compute cost-based distances
oo.dist <- dist(obs.gd$coords)

## compute euclidean (only) variogram
vg.std <- variog(obs.gd)
# str(vg.std)

## fitting variogram models
vgmdl.std <- likfit(geodata = obs.gd,
                    fix.nugget = TRUE,
                    ini = c(10, 5),
                    cov.model = "exponential",
                    lik.method = "REML")

```



```{r variogram-plot, fig.height = 3, fig.cap = 'Empirical variogram and fitted model.'}

ggplot(vg.std) + geom_variogram(vgmdl.std)

```

```{r prediction locations, echo=FALSE}
## Prediction locations
loc <- data.frame(lims)  %>% 
  lapply(function(x) seq(x[1], x[2], by = .05)) %>%
  expand.grid()

## Cost surface
cost_surf <- raster(nrow = 109, ncol = 76,
                    xmn = lims[1, 'X'], xmx = lims[2, 'X'],
                    ymn = lims[1, 'Y'], ymx = lims[2, 'Y'],
                    vals = 1,  # Initial value of "cost"
                    crs = NA)  

cost_loc <- coordinates(cost_surf)
outer.idx <- point.in.polygon(cost_loc[, 'x'],
                              cost_loc[, 'y'],
                              border_coord$x,
                              border_coord$y) == 0
cost_surf[!outer.idx] <- 0   # High friction == low "conductivity"
# image(cost_surf)

## Make sure no one lies within the barrier
## Is this necessary? this complicates plotting
# border <- border_coord%>%
#   Polygon %>%
#   list() %>%
#   Polygons(ID = 'border') %>% 
#   list() %>%
#   SpatialPolygons()
# plot(border)

outer.idx <- point.in.polygon(loc$X,
                              loc$Y,
                              border_coord$x,
                              border_coord$y) == 0
# loc <- loc[outer.idx, ]
```


```{r conventional-kriging, include=FALSE}

# Conventional Kriging, Euclidean distances
KC.std = krige.control(obj.model = vgmdl.std)
kriging.std <- krige.conv(obs.gd, locations = loc, krige = KC.std)
```




```{r cost-based-distances, fig.width=10, fig.height=17, warning=FALSE, echo = FALSE, fig.cap='Cost-based distances to each of the observation locations.'}
## Euclidean distances from data to locations
dl <- apply(coordinates(multielem), 1, dist1, loc)

## Find the locations "behind the wall"
behind.idx <- dist1(origin, loc) <= radius*.95

# ggplot(data.frame(multielem), aes(X, Y)) +
#   geom_point(col = 'lightgray') +
#   geom_polygon(aes(x, y), data = border_coord, col = 'lightgray') +
#   geom_point(aes(X, Y), data = loc, col = as.numeric(behind.idx)+1) +
# #   geom_point(aes(x, y), data = data.frame(coordinates(cost_surf))) +
#   coord_fixed() +
#   theme_bw()

## This is a matrix of "conductivity" between all pairs of neighbouring locations
## as a function of their values, and corrected by the distance
loc_transition <- transition(cost_surf,
                             transitionFunction = min,
                             directions = 16) %>%
  geoCorrection(type = 'c', multpl = FALSE)
# image(transitionMatrix(loc_transition))
# plot(raster(loc_transition))   # mean conductivity wrt neighbors
# summary(transitionMatrix(loc_transition)@x)

## Cost-based distances from prediction to observation locations 
dl.cost <- costDistance(loc_transition, coordinates(loc), coordinates(multielem))

## Cost-based maps to each observation
cb.maps <-  cbind(loc, dl.cost) %>%
  gather("Observation", "Distance", -(X:Y))

ggplot(cb.maps, aes(X, Y)) + 
  geom_tile(aes(fill = Distance, color = Distance)) +
  stat_contour(aes(z = Distance), color = 'lightgray') +
  geom_point(aes(X, Y), data = data.frame(multielem), col = 'white', alpha = .5) +
  geom_polygon(aes(x, y), data = border_coord) +
  scale_fill_gradient(low = '#034E7B', high = '#FDAE6B') +
  scale_color_gradient(low = '#034E7B', high = '#FDAE6B') +
  coord_fixed() +
  facet_wrap(~Observation) 
```



```{r cost-based-kriging, include=FALSE}
## With trend, Cost-based distances
KC.cost = KC.std  # Since the variogram model has not changed

kriging.cost <- krige.conv(obs.gd,
                           locations = loc,
                           krige=KC.cost,
                           #dd.dists.mat = dd.distmat,  # These are Euclidean
                           dl.dists.mat = dl.cost)

```

```{r map-kriging-estimates, fig.width=10, fig.height=7, fig.cap = 'Comparison of Kriging estimates.'}
# image(kriging.std)
# image(kriging.cost)

res.df <- data.frame(loc, 
                     method = c(rep('Classical', nrow(loc)),
                                rep('Cost-based', nrow(loc)),
                                rep('Difference', nrow(loc))),
                     Prediction = c(kriging.std$predict,
                                    kriging.cost$predict,
                                    kriging.std$predict-kriging.cost$predict))

pred.comp <- 
  ggplot(filter(res.df, method != 'Difference'),
         aes(X, Y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_point(aes(X, Y), data = data.frame(multielem), col = 'white', alpha = .5) +
  geom_polygon(aes(x, y), data = border_coord) +
  scale_fill_gradient("Prediction", low = '#2c7fb8', high = '#fdae6b') +
  scale_color_gradient("Prediction", low = '#2c7fb8', high = '#fdae6b') +
  coord_fixed() +
  facet_wrap(~method) +
  theme_bw()

print(pred.comp)
```



```{r map-kriging-difference, fig.width=5, fig.height=7, fig.cap='Difference between the cost-based prediction and the Euclidean prediction'}

pred.dif <- 
  ggplot(filter(res.df, method == 'Difference'),
         aes(X, Y)) + 
  geom_tile(aes(fill = Prediction, colour = Prediction)) +
  stat_contour(aes(z = Prediction), color = 'lightgray') +
  geom_point(aes(X, Y), data = data.frame(multielem), col = 'black', alpha = .5) +
  geom_polygon(aes(x, y), data = border_coord) +
  scale_fill_gradient2(low  ="#3A3A98FF",
                       high = "#832424FF") +
  scale_color_gradient2(low  ="#3A3A98FF",
                       high = "#832424FF") +
  coord_fixed() +
  facet_wrap(~method) +
  theme_bw()

print(pred.dif)
# grid.arrange(pred.comp, pred.dif)
```

