---
title: "Analysis of a Point Pattern in R"
author: "Facundo Muñoz"
date:  "June 03 2015"
output:
  pdf_document:
    fig_caption: true
    pandoc_args: [
      "--output", "../reports/Point_Pattern-faunab10.pdf"
    ]
---

Sample script to show how to analyse a Point Pattern in R.
We model the dataset with an inhomogeneous Poisson Process, estimate its intensity function, and compute its gradient rate.


```{r packages, include = FALSE}
library(spatstat)   # density.ppp(): compute KDE of intensity surface
library(raster)     # terrain(): compute slope of intensity surface
library(ggplot2)    # nicer plots
```

Read coordinates from csv.
Units are meters? Rescale to km.

```{r load}
coord <- read.csv('../data/fauna_b10_R.csv',
                  header = FALSE,
                  col.names = c('x', 'y'))
coord <- coord/1e3
```

```{r remove-duplicates}
coord <- coord[!duplicated(coord),]
```


Take a look at the data. 

```{r plot-coord}
ggplot(coord, aes(x, y)) +
  geom_point() +
  coord_fixed()
```


Create a `spatstat`'s Point Pattern object from the coordinates.

```{r create-ppp}
# Define region as a rectangle containing all the points
# with limits extended by 4%
extend.range <- function(x, pct) {
  range <- range(x)
  range + c(-1, 1) * pct * diff(range) / 100
}
pct = 4
lims <- sapply(coord, extend.range, pct)
dat.ppp <- as.ppp(coord, as.owin(c(lims)))
mean.intensity <- with(dat.ppp, n/diff(window$xrange)/diff(window$yrange))
```

The \emph{mean intensity} is the average density of points over the region, and takes the value
```r mean.intensity``` in this case.


Compute a Kernel Density Estimate of the intensity function.
See `?spatstat::density.ppp` for details

```{r compute-intensity}
intensity <- density(dat.ppp)
# plot(intensity, main = 'Estimate of intensity')
# Nicer:
ggplot(with(intensity, cbind(expand.grid(x = xcol, y = yrow),
                             z = c(t(v)))),
       aes(x, y, fill = z)) +
  geom_raster() + 
  scale_fill_gradient2(low  ="#3A3A98FF",
              high = "#832424FF",
              midpoint = mean.intensity) +
  coord_fixed()
```


Compute the slope of the intensity surface.
As far as I understand, this is what Maximiano Castillejo calls *spatial gradient*.
Here it is important to make sure that `x` and `y` are *projected* coordinates, and not lat/lon.

```{r compute-slope}
# transform the image as a RasterLayer object
# just use the EPSG code to state that coordinates are planar
intens.rst <- raster(intensity, crs = CRS("+init=epsg:3732"))
slope.rst <- terrain(intens.rst)
# plot(slope.rst, main = 'Slope of intensity')
# Nicer:
ggplot(data.frame(coordinates(slope.rst),
                  z = slope.rst@data@values),
       aes(x, y, fill = z)) +
  geom_raster() + 
  scale_fill_gradient(low  ="#034E7B",
              high = "#FDAE6B") +
  coord_fixed()
```

